---
title: "Reliable rejuvenation signal?"
author: "Luke Hayden"
output: html_document
---

#Setup
###Packages
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
#source("https://bioconductor.org/biocLite.R")
#biocLite("DESeq2")
#source("https://bioconductor.org/biocLite.R")
#biocLite("DESeq2")

library(ggplot2)
library(rmarkdown)
library(gplots)
library(RColorBrewer)
library(reshape2)
library(scales)
library(ggbiplot)
library(devtools)
library(NanoStringNorm)
library(gtools)

multiplot <- function(..., plotlist = NULL, file, cols = 1, layout = NULL) {
  require(grid)
  
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  if (is.null(layout)) {
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots == 1) {
    print(plots[[1]])
    
  } else {
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    for (i in 1:numPlots) {
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

###Data Import

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}


gm <- read.csv("gmark.csv")

setwd("~/Documents/nstringnov17/ctg5and6")
ct5and6.raw <- read.markup.RCC()


setwd("~/Documents/nstringnov17/ctg5hi")
ct5and6hi.raw <- read.markup.RCC()


setwd("~/Documents/nstringnov17/ctg7and8")
ct7and8.raw <- read.markup.RCC()


setwd("~/Documents/nstringnov17/ctg7and8hi")
ct7and8hi.raw <- read.markup.RCC()

setwd("~/Documents/nstringnov17/ctg3and4")

ct3and4.raw <- read.markup.RCC()


setwd("~/Documents/nstringnov17/ctg1and2")

ct1and2.raw <- read.markup.RCC()


setwd("~/Documents/nstringnov17/ctg9")

ct9.raw <- read.markup.RCC()

````

###Normalisation

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}
codeset <- read.table("CodeSet.txt")
codeset$shortname <- substr(codeset$Customer_Identifier, 1,2)
Cht <- read.table("CodeSetCountdata.txt")

ctg1and2.norm <- NanoStringNorm(x = ct1and2.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg3and4.norm <- NanoStringNorm(x = ct3and4.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg5and6.norm <- NanoStringNorm(x = ct5and6.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg5and6hi.norm <- NanoStringNorm(x = ct5and6hi.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg7and8.norm <- NanoStringNorm(x = ct7and8.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg7and8hi.norm <- NanoStringNorm(x = ct7and8hi.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);


ctg9.norm <- NanoStringNorm(x = ct9.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);



````

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}

ctg5and6.unnorm <- NanoStringNorm(x = ct5and6.raw, anno = NA, CodeCount ='sum', Background ="mean",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg5and6hi.unnorm <- NanoStringNorm(x = ct5and6hi.raw, anno = NA, CodeCount ='sum', Background ="mean",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg7and8.unnorm <- NanoStringNorm(x = ct7and8.raw, anno = NA, CodeCount ='sum', Background ="mean",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg7and8hi.unnorm <- NanoStringNorm(x = ct7and8hi.raw, anno = NA, CodeCount ='sum', Background ="mean",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg3and4.unnorm <- NanoStringNorm(x = ct3and4.raw, anno = NA, CodeCount ='sum', Background ="mean",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);


ctg1and2.unnorm <- NanoStringNorm(x = ct1and2.raw, anno = NA, CodeCount ='sum', Background ="mean",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg9.unnorm <- NanoStringNorm(x = ct9.raw, anno = NA, CodeCount ='sum', Background ="mean",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg1to9un <- cbind(ctg1and2.unnorm,ctg3and4.unnorm, ctg5and6hi.unnorm, ctg7and8hi.unnorm, ctg9.unnorm)
````

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}


colnames(ctg1and2.norm) <- c("C200", "C100", "C50", "C25", "C12", "C6", "CIA", "CIB", "FYA", "FOA", "MYA", "MOA","FYB",	"FOB",	"MYB",	"MOB",	"MOIA",	"MOIB","MOIC",	"MOID",	"CS1X",	"CS0.5X",	"CS0.25XA",	"CS0.25XB")
colnames(ctg1and2.unnorm) <- colnames(ctg1and2.norm) 

colnames(ctg3and4.norm) <- c("O1a", "O2a", "Y1a", "Y2a", "OR1a", "OR2a", "YR1a", "YR2a", "FY5a", "FY6a", "FO5a", "FO7a", "O1b", "O2b", "Y1b", "Y2b", "OR1b", "OR2b", "YR1b", "YR2b", "FY7", "FY8", "FO8", "FO9")
colnames(ctg3and4.unnorm) <- colnames(ctg3and4.norm) 

colnames(ctg5and6.norm) <- c("SFPA", "SFPB", "LFPA", "LFPB", "SMPA", "SMPB", "LMPA", "LMPB", "LMIA", "LMIB", "LMIC", "LMID", "Abef", "Aun", "Areg", "Bbef", "Bun", "Breg", "Cbef", "Cun", "Creg", "Dbef", "Dun", "Dreg")
colnames(ctg5and6.unnorm) <- colnames(ctg5and6.norm) 

colnames(ctg5and6hi.norm) <-  c("SFPA", "SFPB", "LFPA", "LFPB", "SMPA", "SMPB", "LMPA", "LMPB", "LMIA", "LMIB", "LMIC", "LMID", "Abef", "Aun", "Areg", "Bbef", "Bun", "Breg", "Cbef", "Cun", "Creg", "Dbef", "Dun", "Dreg")
colnames(ctg5and6hi.unnorm) <- colnames(ctg5and6hi.norm) 

colnames(ctg7and8.norm) <- c("SFPC","SFPD", "LFPC", "LFPD", "SMPC", "SMPD","LMPC", "LMPD", "0mPA", "0mPB", "10mPA", "20mPA", "Ebef", "Eun", "Ereg", "Fbef", "Fun", "Freg", "Gbef", "Gun", "Greg", "LFIA", "LFIB", "20mPB")
colnames(ctg7and8.unnorm) <- colnames(ctg7and8.norm) 

colnames(ctg7and8hi.norm) <-c("SFPC","SFPD", "LFPC", "LFPD", "SMPC", "SMPD","LMPC", "LMPD", "0mPA", "0mPB", "10mPA", "20mPA", "Ebef", "Eun", "Ereg", "Fbef", "Fun", "Freg", "Gbef", "Gun", "Greg", "LFIA", "LFIB", "20mPB")
colnames(ctg7and8hi.unnorm) <- colnames(ctg7and8hi.norm) 

colnames(ctg9.norm) <- c("LFIC", "LFID", "OmID", "10mIA", "n", "20mIC", "Ibef", "Iun", "Ireg", "Kbef", "Kun", "Kreg")
colnames(ctg9.unnorm) <- colnames(ctg9.norm) 

ctg3to9 <- cbind(ctg3and4.norm, ctg5and6hi.norm, ctg7and8hi.norm, ctg9.norm)

ctg1to9 <- cbind(ctg1and2.norm,ctg3and4.norm, ctg5and6hi.norm, ctg7and8hi.norm, ctg9.norm)


sampleinfo <- data.frame(

sample=colnames(ctg1to9), 
                         
sex=c(rep("M", 8), rep(c("F", "F", "M", "M"), 2), rep("M", 8), rep("F", 28), rep("M", 8), rep("F", 16), rep("M", 4), rep("F", 28)), 

age=c(rep("O",8), rep(c("Y","O","Y","O"),2),rep("O",8),  rep(c("O", "O", "Y", "Y", "O", "O", "Y", "Y","Y", "Y", "O", "O"), 2), c("Y", "Y", "O", "O", "Y", "Y", "O", "O", "O", "O", "O", "O", rep("O", 12)), c("Y", "Y", "O", "O", "Y", "Y", "O", "O", "N", "N", "N", "N"), rep("O", 11), c("N", "O", "O",  "N", "N", "N", "N", "O", "O", "O", "O", "O", "O")), 

reg=c(rep("",24), rep(c(rep("", 4), rep("R", 4), rep("", 4)), 2), c(rep("", 12), rep(c("", "", "R"),4 )   ), c(rep("", 12), rep(c("", "", "R"),3 )   ), rep("",11 ), "R", "", "", "R") , 

ctg=c(rep(1, 12), rep(2, 12),rep(3, 12), rep(4, 12), rep(5, 12), rep(6, 12), rep(7, 12), rep(8, 12), rep(9, 12)),

size= c(rep("P",6), "I", "I", rep("P",8),rep("I",4), rep("I",4), rep("P", 32), rep("I", 4), rep("P",27), rep("I", 8), "P", rep("I", 4), rep("I", 8)), 

moult = c(rep("", 80), 0,0,10,20,rep("", 11),20,"","", 0,10,"", 20,rep("", 6) ),

 exp= c(rep("conc",8), rep("OvY", 12), rep("cut", 4), rep("OvY",4), rep("regen", 4), rep("OvY",8), rep("regen", 4), rep("OvY",4), rep("OvY", 12), rep("regen", 12), rep("OvY", 8), rep("moult", 4),rep("regen", 9), "OvY", "OvY", "moult",  "OvY", "OvY", "moult", "moult", "moult", "moult", rep("regen", 6) )

)


sampleinfo$flag <- FALSE
sampleinfo$flag[c(5,6,7,8,25,31,32, 54, 59, 66,67, 78,90, 97,101)] <- TRUE


#c(23,24,37,43,44, 54, 59, 66,67, 78,90, 97,101)

sampleinfo$type <- paste(sampleinfo$age, sampleinfo$reg, sep="")
sampleinfo$qual <- "ok"
sampleinfo$qual[c(25 , 31 , 32 , 49,  54 , 67  ,76,  90, 101, 104, 107)] <- "bad"
#sampleinfo$moult <- c(rep(NA, 80), 0,0,10,20,rep(NA, 11),20,NA,NA, 0,10,NA, 20,rep(NA, 6) )

#print("Samples as follows (all experiments)")
#print(paste("males: " sum()))

save(sampleinfo,ctg1to9,ctg1to9intonly,ctg1to9un, codeset, hk.norm, hk.raw, file="nov17.rdata")
save(ctg1to9, file="ctg1to9.rdata")
save(ctg1to9, file="ctg1to9intonly.rdata")
save(ctg1to9un, file="ctg1to9un.rdata")
save(codeset, file="codeset.rdata")


```


#What is our best set of markers?

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
md <- t(subset(t(ctg1to9), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$exp == "OvY" & sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y")& sampleinfo$size== "P" ))
si <- subset(sampleinfo, sampleinfo$sample %in% colnames(md))
old <- t(subset(t(md), si$age == "O"))
young <- t(subset(t(md), si$age == "Y"))
regen <- t(subset(t(ctg1to9), sampleinfo$type == "OR" & sampleinfo$qual == "ok"))

nmd <- as.data.frame(md/rowMeans(md))
nmd$sname <- substr(rownames(md), 1,2)
nmd <- nmd[order(nmd$sname),]
nmd$up <- codeset$log2FoldChange > 0
nmd$YvOl2fc <- foldchange2logratio(foldchange(rowMeans(young), rowMeans(old)))
nmd$ROvOl2fc <- foldchange2logratio(foldchange(rowMeans(regen), rowMeans(old)))

oy <- cbind(old, young)
t.result <- apply(oy, 1, function (x) t.test(x[1:ncol(old)],x[ncol(old)+1:ncol(oy)]))
nmd$p_value <- unlist(lapply(t.result, function(x) x$p.value))


nmd$sname <- factor(nmd$sname, levels=nmd$sname[order(nmd$p_value)])

nmd$dir <- "downregulated with age"
nmd$dir[nmd$up == FALSE] <- "upregulated with age"

nmd$sig <- nmd$p_value <0.05
nmd$sp <- round(nmd$p_value, 2)
nmd$sigstar <- ""
nmd$sigstar[nmd$p_value <0.05] <- "*"

nmd <- nmd[order(nmd$sname),]
codeset <- codeset[codeset$Customer_Identifier,]


nmd$samedir <- sign(codeset$log2FoldChange) == sign(nmd$YvOl2fc) 

mnmd <- melt(nmd, id.vars=c("p_value", "sname", "up", "dir", "sp", "sig", "sigstar", "samedir", "YvOl2fc", "ROvOl2fc"))
mnmd$age <- "old"
mnmd$age[mnmd$variable %in% colnames(young)] <- "young"
mnmd$ing <- ""
mnmd$ing[mnmd$variable %in% gm$marker] <- "."

ggplot(mnmd, aes(x= sname, y=value, colour=age)) +
  theme_bw()+
#  geom_point(size=3, shape=95)+
 stat_summary()+
  scale_y_continuous(trans="log2")+
  facet_grid(~dir, space="free",scales="free" )+
#  geom_boxplot(outlier.shape = NA)+
  scale_colour_manual(values=c("cornflower blue", "red3"))+ 
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=4))+
  geom_text(colour="black", aes(label=sigstar,y=10))+
  geom_text(aes(label=variable), size=1)+
  geom_text(colour="green", aes(label=ing,y=0.015))
  


#  geom_text(colour="black", angle=90, inherit.aes=FALSE, data=nmd, aes(x=sname, y=0, label=sp, alpha=sig))


gmnmd <- subset(mnmd, (mnmd$p_value <0.05 & mnmd$samedir == TRUE))
gmnmd$sname <- as.character(gmnmd$sname)

ggplot(gmnmd, aes(x= sname, y=value, colour=age, label=variable)) +
  theme_bw()+
  geom_point(size=3, shape=95)+
 stat_summary()+
  scale_y_continuous(trans="log2")+
  facet_grid(~dir, space="free",scales="free" )+
#  geom_boxplot(outlier.shape = NA)+
  scale_colour_manual(values=c("cornflower blue", "red3")) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=6))+
  geom_text(size=2)
```



```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
gnmd <- subset(nmd, nmd$p_value <0.05 & nmd$samedir == TRUE)

a=ggplot(data=gnmd, aes(x=YvOl2fc, y=ROvOl2fc, label=sname))+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_bw()+
  xlim(-5,+5)+
  ylim(-5,5)+
 scale_colour_manual(values= c("forest green", "cornflower blue"), name="Marker", labels=c("other","61 chosen"))+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")+
#  geom_rect(inherit.aes=FALSE, xmin=-10, ymin=-10, xmax=0, ymax=0, fill="yellow", alpha=0.004)+
#  geom_rect(inherit.aes=FALSE,xmin=0, ymin=0, xmax=10, ymax=10, fill="yellow", alpha=0.004)+
  geom_point( size=1.5, colour="cornflower blue")+
  geom_abline(inherit.aes=FALSE, intercept=0, slope=1, linetype=2)+
  theme(    plot.background = element_rect(fill = "transparent",colour = "transparent") )+
  coord_fixed(ratio=1)+
  theme(strip.text=element_blank())
  



ggsave(plot=a, height=7.5,width=15,dpi=200, device="png", filename=paste("nplot.png"), limitsize = FALSE,bg = "transparent")



b=ggplot(data=gnmd, aes(x=YvOl2fc, y=0, label=sname))+
  geom_hline(yintercept=0, colour="black")+
  theme_bw()+
  xlim(-5,+5)+
  ylim(-5,5)+
 scale_colour_manual(values= c("forest green", "cornflower blue"), name="Marker", labels=c("other","61 chosen"))+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("")+
#  geom_rect(inherit.aes=FALSE, xmin=-10, ymin=-10, xmax=0, ymax=0, fill="yellow", alpha=0.004)+
#  geom_rect(inherit.aes=FALSE,xmin=0, ymin=0, xmax=10, ymax=10, fill="yellow", alpha=0.004)+
  geom_point( size=1.5, colour="cornflower blue")+
  theme(    plot.background = element_rect(fill = "transparent",colour = "transparent") )+
  coord_fixed(ratio=1)+
  theme(strip.text=element_blank())
 

ggsave(plot=b, height=7.5,width=15,dpi=200, device="png", filename=paste("nplotb.png"), limitsize = FALSE,bg = "transparent")
 

c=ggplot(data=gnmd, aes(x=YvOl2fc, y=0, label=sname))+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_bw()+
  xlim(-5,+5)+
  ylim(-5,5)+
 scale_colour_manual(values= c("forest green", "cornflower blue"), name="Marker", labels=c("other","61 chosen"))+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")+
#  geom_rect(inherit.aes=FALSE, xmin=-10, ymin=-10, xmax=0, ymax=0, fill="yellow", alpha=0.004)+
#  geom_rect(inherit.aes=FALSE,xmin=0, ymin=0, xmax=10, ymax=10, fill="yellow", alpha=0.004)+
  geom_point( size=1.5, colour="cornflower blue")+
  geom_abline(inherit.aes=FALSE, intercept=0, slope=1, linetype=2)+
  theme(    plot.background = element_rect(fill = "transparent",colour = "transparent") )+
  coord_fixed(ratio=1)+
  theme(strip.text=element_blank())


ggsave(plot=c, height=7.5,width=15,dpi=200, device="png", filename=paste("nplotc.png"), limitsize = FALSE,bg = "transparent")

```



#Nikos plot
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
md <- ctg1to9
old <- t(subset(t(md), sampleinfo$type =="O" &  sampleinfo$sex == "F" & sampleinfo$qual =="ok" &sampleinfo$ctg %in% c(3,4,5,7)))

young <- t(subset(t(md), sampleinfo$type =="Y" & sampleinfo$sex== "F" & sampleinfo$qual =="ok"&sampleinfo$ctg %in% c(3,4,5,7)))

regen <- t(subset(t(md), sampleinfo$type =="OR" & sampleinfo$qual =="ok" &sampleinfo$ctg>2))

nmd <- md/rowMeans(cbind(young, old))

nik<- data.frame(
  fullname= rownames(md), 
  sname = substr(rownames(md),1,2), 
  YvOl2fc= foldchange2logratio(foldchange(rowMeans(young),            rowMeans(old))),
  ROvOl2fc= foldchange2logratio(foldchange(rowMeans(regen), rowMeans(old))),
  in61 =rownames(md) %in% gm$marker)

snik=subset(nik, nik$in61==TRUE)

ggplot(data=nik, aes(x=YvOl2fc, y=ROvOl2fc, label=sname, colour=in61))+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_bw()+
  xlim(-6,+6)+
  ylim(-6,6)+
 scale_colour_manual(values= c("forest green", "cornflower blue"), name="Marker", labels=c("other","61 chosen"))+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")+
#  geom_rect(inherit.aes=FALSE, xmin=-10, ymin=-10, xmax=0, ymax=0, fill="yellow", alpha=0.004)+
#  geom_rect(inherit.aes=FALSE,xmin=0, ymin=0, xmax=10, ymax=10, fill="yellow", alpha=0.004)+
  geom_point( size=8, alpha=0.4)+
  geom_abline(inherit.aes=FALSE, intercept=0, slope=1, linetype=2)+
  theme(    plot.background = element_rect(fill = "transparent",colour = "transparent") )+
  geom_text()+
  coord_fixed(ratio=1)+
  facet_wrap(~in61)+
  theme(strip.text=element_blank())
  



ggsave(plot=a, height=7.5,width=15,dpi=200, device="png", filename=paste("nplot.png"), limitsize = FALSE,bg = "transparent")


```

#Variation due to Regeneration
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
md <- t(subset(t(ctg1to9), sampleinfo$sex == "F"& sampleinfo$qual == "ok"&  sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y", "OR") ))
si <- subset(sampleinfo, sampleinfo$sample %in% colnames(md))
nmd <- as.data.frame(md/rowMeans(md))
regonly <- t(subset(t(nmd), si$type == "OR"))
nmd$regstdev <- apply(regonly, 1,sd)
nmd <- subset(nmd, rownames(nmd)%in% gm$marker)
nmd$marker <- rownames(nmd)
nmd <- nmd[order(nmd$marker),]
gm <- gm[order(gm$marker),]
nmd$dir <- gm$dir[gm$dir != "housekeeping"]

nmd$smarker <- substr(nmd$marker,1,2)
nmd$smarker <- factor(nmd$smarker, levels=nmd$smarker[order(nmd$regstdev)])


mnmd <- melt(nmd, id.vars=c("marker", "dir", "regstdev", "smarker"))
mnmd$type <- "Y"
mnmd$type[mnmd$variable %in% si$sample[si$type== "OR"]] <- "OR"
mnmd$type[mnmd$variable %in% si$sample[si$type== "O"]] <- "O"


ggplot(mnmd, aes(x= smarker, y=value, colour=type)) +
  theme_bw()+
  scale_y_continuous(trans="log2")+
  facet_grid(~dir, space="free",scales="free" )+
  geom_boxplot(outlier.shape = NA)+
  scale_colour_manual(values=c("cornflower blue", "forest green", "red3" ))+ 
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=6))
#  geom_text(aes(label=variable), size=1)


```

#Reliable rejuvenation?

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
md <- ctg1to9
old <- t(subset(t(md), sampleinfo$type =="O" &  sampleinfo$sex == "F" & sampleinfo$qual =="ok" &sampleinfo$ctg %in% c(3,4,5,7)))

young <- t(subset(t(md), sampleinfo$type =="Y" & sampleinfo$sex== "F" & sampleinfo$qual =="ok"&sampleinfo$ctg %in% c(3,4,5,7)))

regen <- t(subset(t(md), sampleinfo$type =="OR" & sampleinfo$qual =="ok" &sampleinfo$ctg>2))

oy <- cbind(old, young)

t.result <- apply(oy, 1, function (x)
t.test(x[1:ncol(old)],x[ncol(old)+1:ncol(oy)]))
ovy.p_value <- unlist(lapply(t.result, function(x) x$p.value))


oreg <- cbind(old, regen)

t.result <- apply(oreg, 1, function (x)
t.test(x[1:ncol(old)],x[ncol(old)+1:ncol(oreg)]))
ovreg.p_value <- unlist(lapply(t.result, function(x) x$p.value))

nold <- old/rowMeans(md)
nyoung <- young/rowMeans(md)
nregen <- regen/rowMeans(md)

nmd <- md/rowMeans(cbind(young, old))



nik<- data.frame(
  fullname= rownames(md), 
  sname = substr(rownames(md),1,2), 
  YvOl2fc= foldchange2logratio(foldchange(rowMeans(young),            rowMeans(old))),
  ROvOl2fc= foldchange2logratio(foldchange(rowMeans(regen), rowMeans(old))),
  in61 =rownames(md) %in% gm$marker, 
  ovypval = ovy.p_value, 
  ovrpval = ovreg.p_value,
  oldsd=  apply(nold, 1,sd),
  youngsd = apply(nyoung, 1,sd),
  regsd = apply(nregen, 1,sd))

nik$ovysig <- nik$ovypval <0.05
nik$ovrsig <- nik$ovrpval <0.1

ggplot(data=nik, aes(x=YvOl2fc, y=ROvOl2fc, label=sname, fill=regsd, shape=ovrsig))+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_minimal()+
  xlim(-6,+6)+
  ylim(-6,6)+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")+
#  geom_rect(inherit.aes=FALSE, xmin=-10, ymin=-10, xmax=0, ymax=0, fill="yellow", alpha=0.004)+
#  geom_rect(inherit.aes=FALSE,xmin=0, ymin=0, xmax=10, ymax=10, fill="yellow", alpha=0.004)+
  geom_point( size=3)+
  scale_shape_manual(values=c(21,23), labels=c(">0.1", "<0.1"), breaks=c(FALSE, TRUE) ,name="RO vs O p-value" )+
  geom_abline(inherit.aes=FALSE, intercept=0, slope=1, linetype=2)+
  theme(    plot.background = element_rect(fill = "transparent",colour = "transparent") )+
  coord_fixed(ratio=1)+
scale_fill_gradient2(high="red3", mid="white", low="blue", midpoint=4, name="Stdev (regen)")+
  facet_wrap(~ovysig)


nik$good <- nik$ROvOl2fc/nik$YvOl2fc > 0.6 & nik$ovysig ==TRUE &nik$regsd <3
nik$reju <- " neither"
nik$reju[nik$ROvOl2fc/nik$YvOl2fc > 0.6 & nik$ovysig ==TRUE &nik$regsd <3] <- "rejuvenated"
nik$reju[nik$ROvOl2fc/nik$YvOl2fc < -0.6 & nik$ovysig ==TRUE &nik$regsd <3] <- "aged"

ggplot(data=nik, aes(x=YvOl2fc, y=ROvOl2fc, label=sname, fill=regsd,colour=reju, shape=ovrsig))+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_minimal()+
  xlim(-6,+6)+
  ylim(-6,6)+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")+
#  geom_rect(inherit.aes=FALSE, xmin=-10, ymin=-10, xmax=0, ymax=0, fill="yellow", alpha=0.004)+
#  geom_rect(inherit.aes=FALSE,xmin=0, ymin=0, xmax=10, ymax=10, fill="yellow", alpha=0.004)+
  geom_point( size=3.5)+
  scale_shape_manual(values=c(21,23), labels=c(">0.1", "<0.1"), breaks=c(FALSE, TRUE) ,name="Regen vs Old \np-value" )+
  geom_abline(inherit.aes=FALSE, intercept=0, slope=1, linetype=2)+
  theme(    plot.background = element_rect(fill = "transparent",colour = "transparent") )+
  coord_fixed(ratio=1)+
scale_fill_gradient2(high="darkgoldenrod1", mid="white", low="blue", midpoint=2.5, name="Standard deviation \n amongst regenerated")+
  facet_wrap(~ovysig)+
  geom_text(size=1.7, aes(alpha=reju))+
  scale_colour_manual( values=c("red2",   "black", "dark green"), name="Rejuvenation")+
  scale_alpha_manual(values=c(1,0,1))




```


#Rejuventating Markers
Only those with:
Ratio of YvO l2fc / RO vs O l2fc > 0.6
Stdev between regenerated <3

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
md <- t(subset(t(ctg1to9), sampleinfo$sex == "F"& sampleinfo$qual == "ok"&  sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y", "OR") ))
si <- subset(sampleinfo, sampleinfo$sample %in% colnames(md))
nmd <- as.data.frame(md/rowMeans(md))
regonly <- t(subset(t(nmd), si$type == "OR"))
nmd$regstdev <- apply(regonly, 1,sd)

nmd$marker <- rownames(nmd)
nmd <- nmd[order(nmd$marker),]
codeset <- codeset[order(codeset$shortname),]
nmd$dir <- "upregulated with age"
nmd$dir[codeset$log2FoldChange > 0] <- "downregulated with age" 



nmd$smarker <- substr(nmd$marker,1,2)
nmd$smarker <- factor(nmd$smarker, levels=nmd$smarker[order(nmd$regstdev)])
nmd <- subset(nmd, nmd$marker %in% nik$fullname[nik$good ==T])

mnmd <- melt(nmd, id.vars=c("marker", "regstdev", "dir", "smarker"))
mnmd$type <- "Y"
mnmd$type[mnmd$variable %in% si$sample[si$type== "OR"]] <- "OR"
mnmd$type[mnmd$variable %in% si$sample[si$type== "O"]] <- "O"

mcodeset <- subset(codeset, codeset$Customer_Identifier %in% nmd$marker)

print(paste(mcodeset$shortname, ": " ,mcodeset$DrosoTopHit, sep=""))

print(paste(mcodeset$shortname, ": " ,mcodeset$TopHit, sep=""))

mnmd$topblast <- ""

for (i in 1:length(codeset$baseMean)){
  x = mnmd$marker == codeset$Customer_Identifier[i]
  mnmd$topblast[x] <- as.character(codeset$TopHit[i])
  mnmd$dtopblast[x] <- as.character(codeset$DrosoTopHit[i])
}



```



```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

ggplot(mnmd, aes(x= smarker, y=value, fill=type)) +
  theme_minimal()+
  scale_y_continuous(trans="log2")+
  facet_grid(~dir, space="free",scales="free" )+
#  stat_summary()+
  geom_boxplot(outlier.shape = NA)+
#  geom_point(shape=21)+
  scale_fill_manual(values=c("cornflower blue", "forest green", "red3" ))+ 
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=10))+
  xlab("Marker")+
  ylab("Expression level (normalised)")
  
#  geom_text(aes(label=variable), size=1)
mnmd$smarkerdir <- paste(substr(mnmd$dir,1,2), mnmd$smarker)
umnmd <- subset(mnmd, !duplicated(mnmd$marker))



#mcodeset$smarkerdir <- paste(substr(mcodeset$dir,1,2), mnmd$smarker)
  
  
ggplot(mnmd, aes(x= value, fill=type)) +

  theme_minimal()+
  scale_x_continuous(trans="log2", breaks=c(0.125,0.25,0.5,1,2,4), limits=c(0.125,4))+
  facet_wrap(~smarkerdir, ncol=3, scales="free_y")+
  geom_density(alpha=0.5)  +
  scale_fill_manual(values=c("cornflower blue", "forest green", "red3" ))+ 
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=10))+
  xlab("Expression level (normalised)")+
  geom_text(data=umnmd, aes(x=0.2, y=0.6, label=dtopblast), size=2,inherit.aes = F)



```



