---
title: "Reliable rejuvenation signal?"
author: "Luke Hayden"
output: html_document
---

#Setup
###Packages
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
#source("https://bioconductor.org/biocLite.R")
#biocLite("DESeq2")
#source("https://bioconductor.org/biocLite.R")
#biocLite("DESeq2")

library(ggplot2)
library(rmarkdown)
library(gplots)
library(RColorBrewer)
library(reshape2)
library(scales)
library(ggbiplot)
library(devtools)
library(NanoStringNorm)
library(gtools)
library(Rmisc)
library(dplyr)
library(tidyr)
library(caret)
library(e1071)
library(caTools)
library(FinCal)

```

###Data Import

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}
setwd("~/Documents/nstringnov17")
load(file="nov17main_data.rdata")
```

#What is our best set of markers?


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

md <- t(subset(t(ctg1to9), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$exp == "OvY" & sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y")& sampleinfo$size== "P" ))

si <- subset(sampleinfo, sampleinfo$sample %in% colnames(md))
old <- t(subset(t(md), si$age == "O"))
young <- t(subset(t(md), si$age == "Y"))
regen <- t(subset(t(ctg1to9), sampleinfo$type == "OR" & sampleinfo$qual == "ok"))





nmd <- as.data.frame(md/rowMeans(md))
nmd$name <- rownames(nmd)
nmd$sname <- substr(nmd$name, 1,2)

oy <- cbind(old, young)
t.result <-  apply(oy, 1, function (x) t.test(x[1:ncol(old)],x[ncol(old)+1: ncol(oy)]))


nmd$p_value <- unlist(lapply(t.result, function(x) x$p.value))

nmd <- left_join(nmd, geneinf, by="name")



nmd$YvOl2fc <- foldchange2logratio(foldchange(rowMeans(young), rowMeans(old)))
nmd$ROvOl2fc <- foldchange2logratio(foldchange(rowMeans(regen), rowMeans(old)))

nmd$chosen <- nmd$p_value < 0.05 & sign(nmd$transl2fc) == sign(nmd$YvOl2fc)


ggplot(nmd, aes(y=p_value, x=abs(YvOl2fc), colour=in61,shape=in61,  fill=chosen))  +
  scale_shape_manual(values=c(21,22))+
  geom_point()+ 
  theme_minimal()+
  scale_fill_manual(values=c("cornflower blue", "forestgreen"), name="Amongst new chosen set") +
  scale_colour_manual(values=c("red3", "black"))+
  xlab("Absolute Young vs Old Log2Foldchange") +
  ylab("Old vs young p-value")+
  ggtitle(paste(sum(nmd$chosen), " markers: new optimised set"))+
    geom_hline(linetype=2, yintercept = 0.05)
  



```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
nmd$sname <- factor(nmd$sname, levels=nmd$sname[order(nmd$p_value)])

mnmd <- gather(nmd, key=sample, value=normexp, O2a:LFPC)
mnmd$age <- ifelse(mnmd$sample %in% colnames(young), "young", "old")


ggplot(mnmd, aes(x= sname, y=normexp, colour=age)) +
  theme_minimal()+
  geom_point(size=3, shape=95)+
 stat_summary()+
  scale_y_continuous(trans="log2")+
  facet_grid(~dir, space="free",scales="free" )+
#  geom_boxplot(outlier.shape = NA)+
  scale_colour_manual(values=c("cornflower blue", "red3"))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=4))+
  geom_text(colour="black",label="*", data=nmd,inherit.aes=FALSE,  aes(alpha=chosen, x=sname, y=8))+
  scale_alpha_manual(values=c(0,1))


```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}


ggplot(mnmd, aes(x= sname, y=normexp, colour=age)) +
  theme_minimal()+
  geom_boxplot(outlier.shape = NA)+
  scale_y_continuous(trans="log2")+
  facet_grid(~dir, space="free",scales="free" )+
#  geom_boxplot(outlier.shape = NA)+
  scale_colour_manual(values=c("cornflower blue", "red3"))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=4))+
  geom_text(colour="black",label="*", data=nmd,inherit.aes=FALSE,  aes(alpha=chosen, x=sname, y=8))


```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
minf <- nmd[,colnames(nmd) %in% c("name"   ,   "sname"    , "p_value",   "transl2fc", "dir"  ,     "in61"    ,  "YvOl2fc"  , "ROvOl2fc",  "chosen" )]





```


#Compare old vs new marker set

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

log <- sampleinfo$sex =="F" & sampleinfo$ctg %in% c(3,4,5,7) & sampleinfo$qual == "ok" & sampleinfo$reg == "" & sampleinfo$age != "N"

bcmin <- ctg1to9[,log== TRUE]
bcmin.groups <- subset(sampleinfo$type, log==TRUE)
bcmin <- subset(bcmin, rownames(bcmin) %in% geneinf$name[geneinf$in61==T])
tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups


a=ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=FALSE, labels=rownames(tbc), groups=bc.groups)+ggtitle("PCA of Old vs Young (females): 61 chosen markers")+
    theme_minimal() +
  scale_colour_manual(values=c("cornflower blue", "red3", "black"))+
  theme(plot.title=element_text(size=8,face="bold"))



bcmin <- ctg1to9[,log== TRUE]
bcmin.groups <- subset(sampleinfo$type, log==TRUE)
bcmin <- subset(bcmin, rownames(bcmin) %in% nmd$name[nmd$chosen==TRUE])
tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups

b=ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=FALSE, labels=rownames(tbc), groups=bc.groups)+ggtitle(paste("PCA of Old vs Young (females):", nrow(bcmin) ," newly chosen markers"))+
    theme_minimal() +
  scale_colour_manual(values=c("cornflower blue", "red3", "black"))+
  theme(plot.title=element_text(size=8,face="bold"))


multiplot(a,b, cols=2)


```

Without Nikos' samples:

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

log <- sampleinfo$sex =="F" & sampleinfo$ctg %in% c(3,4,5,7) & sampleinfo$qual == "ok" & sampleinfo$reg == "" & sampleinfo$age != "N" &sampleinfo$prep == "Luke"

bcmin <- ctg1to9[,log== TRUE]
bcmin.groups <- subset(sampleinfo$type, log==TRUE)
bcmin <- subset(bcmin, rownames(bcmin) %in% geneinf$name[geneinf$in61==T])
tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups


a=ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=FALSE, labels=rownames(tbc), groups=bc.groups)+ggtitle("PCA of Old vs Young (females): 61 chosen markers")+
    theme_minimal() +
  scale_colour_manual(values=c("cornflower blue", "red3", "black"))+
  theme(plot.title=element_text(size=8,face="bold"))



bcmin <- ctg1to9[,log== TRUE]
bcmin.groups <- subset(sampleinfo$type, log==TRUE)
bcmin <- subset(bcmin, rownames(bcmin) %in% nmd$name[nmd$chosen==TRUE])
tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups

b=ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=FALSE, labels=rownames(tbc), groups=bc.groups)+ggtitle(paste("PCA of Old vs Young (females):", nrow(bcmin) ,"newly chosen markers"))+
    theme_minimal() +
  scale_colour_manual(values=c("cornflower blue", "red3", "black"))+
  theme(plot.title=element_text(size=8,face="bold"))


multiplot(a,b, cols=2)


```


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
gnmd <- subset(nmd, nmd$chosen== TRUE)

ggplot(data=nmd, aes(x=YvOl2fc, y=ROvOl2fc, label=sname, colour=chosen))+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_minimal()+
  xlim(-5,+5)+
  ylim(-5,5)+
 scale_colour_manual(values= c("grey", "forest green"), name="Marker", labels=c("other","chosen"))+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")+
#  geom_rect(inherit.aes=FALSE, xmin=-10, ymin=-10, xmax=0, ymax=0, fill="yellow", alpha=0.004)+
#  geom_rect(inherit.aes=FALSE,xmin=0, ymin=0, xmax=10, ymax=10, fill="yellow", alpha=0.004)+
  geom_point( size=1.5)+


  geom_abline(inherit.aes=FALSE, intercept=0, slope=1, linetype=2)+
  theme(    plot.background = element_rect(fill = "transparent",colour = "transparent") )+
  coord_fixed(ratio=1)
  

```

#Logistic regression model to separate old from young

First: use all markers as predictors of status in old vs young

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

dmat<- as.data.frame(subset(t(ctg1to9), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$exp == "OvY" & sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y")& sampleinfo$size== "P" ))


dmat$age <- rownames(dmat) %in% colnames(young)

#logmod <- glm(age~., dmat, family='binomial')
logmod <- lm(age~., dmat,method="binomial" )

pred <- predict(logmod, type='response')

predclass <- ifelse(pred> 0.5, T, F)

confusionMatrix(predclass, dmat$age)

colAUC(pred, dmat$age, plotROC=T)


(modcoefs <- data.frame(logmodcoef=coef(logmod), name=names(coef(logmod))))


minf2 <- left_join(minf, modcoefs, by="name")

```


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

dmat$age <- ifelse(rownames(dmat) %in% colnames(young), "young", "old")

myControl <- trainControl(
  method = "cv",
  number = 4,
  summaryFunction = twoClassSummary,
  classProbs = TRUE, # IMPORTANT!
  verboseIter = TRUE
)

(logmodcv <- train(age~.,dmat,method='glm', trControl=myControl ))

#ggplot( cbind(dmat, pred), aes(x=age, y=pred))+geom_boxplot()

predcv <- predict(logmodcv)




(modcoefs <- data.frame(logmodcoef=coef(logmodcv$finalModel), name=names(coef(logmodcv$finalModel))))


minf2 <- left_join(minf, modcoefs, by="name")

```

#Nikos plot
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
md <- ctg1to9

old <- t(subset(t(md), sampleinfo$type =="O" &  sampleinfo$sex == "F" & sampleinfo$qual =="ok" &sampleinfo$ctg %in% c(3,4,5,7)))

young <- t(subset(t(md), sampleinfo$type =="Y" & sampleinfo$sex== "F" & sampleinfo$qual =="ok"&sampleinfo$ctg %in% c(3,4,5,7)))

regen <- t(subset(t(md), sampleinfo$type =="OR" & sampleinfo$qual =="ok" &sampleinfo$ctg>2))

vardat <- data.frame(name=rownames(old), 
                     oldmean=rowMeans(old), 
                     oldsd= apply(old,1, sd), 
                     oldcv = coefficient.variation(rowMeans(old), apply(old,1, sd)),
                   youngmean=rowMeans(young), 
                     youngsd= apply(young,1, sd), 
                     youngcv = coefficient.variation(rowMeans(young), apply(young,1, sd)), 
                   regenmean=rowMeans(regen), 
                     regensd= apply(regen,1, sd), 
                     regencv = coefficient.variation(rowMeans(regen), apply(regen,1, sd))
                   ) %>%
  left_join(minf, by="name")


oreg <- cbind(old, regen)

t.result <- apply(oreg, 1, function (x)
t.test(x[1:ncol(old)],x[ncol(old)+1:ncol(oreg)]))
vardat$ovreg.pval <- unlist(lapply(t.result, function(x) x$p.value))
vardat$ovregsig <- vardat$ovreg.pval <0.1

ggplot(data=vardat, aes(x=YvOl2fc, y=ROvOl2fc, label=sname, colour=chosen))+
  geom_point()+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_bw()+
  xlim(-6,+6)+
  ylim(-6,6)+
  scale_colour_manual(values= c("grey", "forest green"), name="Marker", labels=c("other","chosen"))+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")


```



```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
vardat2 <- subset(vardat, vardat$chosen ==TRUE)

ggplot(data=vardat2, aes(x=YvOl2fc, y=ROvOl2fc, label=sname, fill=regencv, shape=ovregsig))+
  geom_point(size=3)+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_minimal()+
  xlim(-6,+6)+
  ylim(-6,6)+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")+
  scale_shape_manual(values=c(21,23), labels=c(">0.1", "<0.1"), breaks=c(FALSE, TRUE) ,name="RO vs O p-value" )+
  scale_fill_gradient2(high="red3", mid="white", low="blue", midpoint=0.6, name="cv (regen)")+
  geom_abline(inherit.aes=FALSE, intercept=0, slope=1, linetype=2)



```


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

ggplot(data=vardat2, aes(y=ROvOl2fc, label=sname, x=regencv, shape=ovregsig, fill=dir))+
  geom_point(size=3)+
  theme_minimal()+
  xlab("Coefficient of variation (amongst regenerated samples)") +
 ylab("Regenerated vs Old Log2 Fold Change")+
  scale_fill_manual(values=c("dark blue", "forest green"))+
 scale_shape_manual(values=c(21,23), labels=c(">0.1", "<0.1"), breaks=c(FALSE, TRUE) ,name="RO vs O p-value" )

```

#Rejuvenating Markers
Only those with:
from chosen 80 markers
Ratio of YvO l2fc / RO vs O l2fc > 0.6
cv between regenerated <0.5

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
vardat$rejuv <- "other"


vardat$rejuv[vardat$chosen== TRUE & vardat$regencv < 0.5 & vardat$YvOl2fc/vardat$ROvOl2fc > 0.6] <- "rejuvenated"


vardat$rejuv[vardat$chosen== TRUE & vardat$regencv < 0.5 & vardat$YvOl2fc/vardat$ROvOl2fc < -0.6] <- "aged"


vardat2 <- subset(vardat, vardat$chosen ==TRUE)
ggplot(data=vardat2, aes(x=YvOl2fc, y=ROvOl2fc, label=sname, fill=regencv, shape=ovregsig, colour=rejuv))+
  geom_point(size=3)+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_minimal()+
  xlim(-6,+6)+
  ylim(-6,6)+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")+
  scale_shape_manual(values=c(21,23), labels=c(">0.1", "<0.1"), breaks=c(FALSE, TRUE) ,name="RO vs O p-value" )+
  scale_fill_gradient2(high="red3", mid="white", low="blue", midpoint=0.6, name="cv (regen)")+
  geom_abline(inherit.aes=FALSE, intercept=0, slope=1, linetype=2)



#vardat$antirejuv <-  vardat$chosen== TRUE & vardat$regencv < 0.5 & vardat$YvOl2fc/vardat$ROvOl2fc > 0.6

```


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
md <- ctg1to9[,colnames(ctg1to9) %in% c(colnames(old), colnames(young), colnames(regen))]
md <- data.frame(md/rowMeans(md)) 
md$name <- rownames(md)
md <- inner_join(md, vardat2, by="name") %>% 
  gather(key="sample", value="normexp", 1:sum(ncol(old), ncol(young), ncol(regen))) %>%
  left_join(sampleinfo, by="sample")%>%  inner_join(codeset, by=c("name" = "Customer_Identifier"))

ggplot(data=subset(md, md$rejuv != "other"), aes(x=sname, y=normexp, fill=type))+
#  geom_point(shape=95)+
  geom_boxplot(outlier.shape = NA)+
  geom_hline(yintercept=1, linetype=2)+
#  stat_summary()+
  facet_grid(~rejuv, space="free", scale="free")+
  scale_fill_manual(values=c("cornflower blue", "forest green", "red3"))+
  theme_minimal()+
  scale_y_continuous(trans="log2", breaks=c(0.125,0.25,0.5,1,2,4))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=10))+
  xlab("Marker")+
  ylab("Expression level (normalised)")

print("rejuvenated:")
levels(as.factor(as.character(subset(md, md$rejuv == "rejuvenated")$TopHit)))

print("aged:")
levels(as.factor(as.character(subset(md, md$rejuv == "aged")$TopHit)))

```

