---
title: "Reliable rejuvenation signal?"
author: "Luke Hayden"
output: html_document
---

#Setup
###Packages
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
#source("https://bioconductor.org/biocLite.R")
#biocLite("DESeq2")
#source("https://bioconductor.org/biocLite.R")
#biocLite("DESeq2")

library(ggplot2)
library(rmarkdown)
library(gplots)
library(RColorBrewer)
library(reshape2)
library(scales)
library(ggbiplot)
library(devtools)
library(NanoStringNorm)
library(gtools)
library(Rmisc)
```

###Data Import

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}
setwd("~/Documents/nstringnov17")
load(file="nov17main_data.rdata")
```

#What is our best set of markers?

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
md <- t(subset(t(ctg1to9), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$ctg %in% c(3,4) & sampleinfo$type %in% c("O", "Y")))
si <- subset(sampleinfo, sampleinfo$sample %in% colnames(md))
old <- t(subset(t(md), si$age == "O"))
young <- t(subset(t(md), si$age == "Y"))
regen <- t(subset(t(ctg1to9), sampleinfo$type == "OR" & sampleinfo$qual == "ok"))

nmd <- as.data.frame(md/rowMeans(md))


nmd$sname <- substr(rownames(md), 1,2)
nmd$name <- rownames(md)
nmd <- arrange(nmd, name)
minf <- arrange(minf, name)


nmd <- nmd[order(nmd$sname),]
nmd$dir <- ifelse(nmd$name %in% minf$name[minf$alldir == "downregulated with age"], "downregulated with age", "upregulated with age")

#nmd$dir <- "upregulated with age"
#nmd$dir[nmd$name %in% minf$name[minf$alldir == "downregulated with age"]] <- "downregulated with age"


nmd$YvOl2fc <- foldchange2logratio(foldchange(rowMeans(young), rowMeans(old)))
nmd$ROvOl2fc <- foldchange2logratio(foldchange(rowMeans(regen), rowMeans(old)))

oy <- cbind(old, young)
t.result <- apply(oy, 1, function (x) t.test(x[1:ncol(old)],x[ncol(old)+1:ncol(oy)]))
nmd$p_value <- unlist(lapply(t.result, function(x) x$p.value))




plot( abs(nmd$YvOl2fc), nmd$p_value)


#ggplot(nmd, aes(x=p_value, y=YvOl2fc))+  geom_point()+  scale_x_log10()

nmd$sname <- factor(nmd$sname, levels=nmd$sname[order(nmd$p_value)])

nmd$in61 <- nmd$name %in% minf$name[minf$in61]

nmd$sig <- nmd$p_value <0.05
nmd$sp <- round(nmd$p_value, 2)
nmd$sigstar <- ""
nmd$sigstar[nmd$p_value <0.05] <- "*"
nmd$in61star <- ""
nmd$in61star[nmd$in61 == TRUE] <- "*"



ggplot(nmd, aes(y=p_value, x=abs(YvOl2fc), colour=in61))  +geom_point()



```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

nmd <- nmd[order(nmd$sname),]
codeset <- codeset[codeset$Customer_Identifier,]


nmd$samedir <- sign(codeset$log2FoldChange) == sign(nmd$YvOl2fc) 

mnmd <- melt(nmd, id.vars=c("p_value", "sname","dir", "sp", "sig", "sigstar", "samedir", "YvOl2fc", "ROvOl2fc", "in61", "in61star", "name"))
mnmd$age <- "old"
mnmd$age[mnmd$variable %in% colnames(young)] <- "young"
mnmd$ing <- ""
mnmd$ing[mnmd$variable %in% gm$marker] <- "."

ggplot(mnmd, aes(x= sname, y=value, colour=age)) +
  theme_bw()+
  geom_point(size=3, shape=95)+
 stat_summary()+
  scale_y_continuous(trans="log2")+
  facet_grid(~dir, space="free",scales="free" )+
#  geom_boxplot(outlier.shape = NA)+
  scale_colour_manual(values=c("cornflower blue", "red3"))+ 
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=4))+
  geom_text(colour="black", aes(label=sigstar,y=10))+
  geom_text(colour="green", aes(label=ing,y=0.015))+
  geom_text(colour="forest green", aes(label=in61star,y=0.1))

sum(nmd$sig)


ggplot(nmd, aes(x=YvOl2fc, y=p_value, colour=in61))+geom_point()
  

```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
#  geom_text(colour="black", angle=90, inherit.aes=FALSE, data=nmd, aes(x=sname, y=0, label=sp, alpha=sig))


gmnmd <- subset(mnmd, (mnmd$p_value <0.05 & mnmd$samedir == TRUE))
gmnmd$sname <- as.character(gmnmd$sname)

ggplot(gmnmd, aes(x= sname, y=value, colour=age, label=variable)) +
  theme_bw()+
  geom_point(size=3, shape=95)+
 stat_summary()+
  scale_y_continuous(trans="log2")+
  facet_grid(~dir, space="free",scales="free" )+
#  geom_boxplot(outlier.shape = NA)+
  scale_colour_manual(values=c("cornflower blue", "red3")) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=6))+
  geom_text(size=2)
```


#Further optimisation of marker set
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
md <- t(subset(t(ctg1to9), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$exp == "OvY" & sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y")& sampleinfo$size== "P" ))
si <- subset(sampleinfo, sampleinfo$sample %in% colnames(md))
old <- t(subset(t(md), si$age == "O"))
young <- t(subset(t(md), si$age == "Y"))
regen <- t(subset(t(ctg1to9), sampleinfo$type == "OR" & sampleinfo$qual == "ok"))

nmd <- as.data.frame(md/rowMeans(md))
nmd$sname <- substr(rownames(md), 1,2)
nmd <- nmd[order(nmd$sname),]
nmd$up <- codeset$log2FoldChange > 0
nmd$YvOl2fc <- foldchange2logratio(foldchange(rowMeans(young), rowMeans(old)))
nmd$ROvOl2fc <- foldchange2logratio(foldchange(rowMeans(regen), rowMeans(old)))

oy <- cbind(old, young)
t.result <- apply(oy, 1, function (x) t.test(x[1:ncol(old)],x[ncol(old)+1:ncol(oy)]))
nmd$p_value <- unlist(lapply(t.result, function(x) x$p.value))


nmd$sname <- factor(nmd$sname, levels=nmd$sname[order(nmd$p_value)])

nmd$dir <- "downregulated with age"
nmd$dir[nmd$up == FALSE] <- "upregulated with age"

nmd$sig <- nmd$p_value <0.05
nmd$sp <- round(nmd$p_value, 2)
nmd$sigstar <- ""
nmd$sigstar[nmd$p_value <0.05] <- "*"

nmd <- nmd[order(nmd$sname),]
codeset <- codeset[codeset$Customer_Identifier,]


nmd$samedir <- sign(codeset$log2FoldChange) == sign(nmd$YvOl2fc) 


ngm <- rownames(nmd)[nmd$p_value <0.05 & nmd$samedir == TRUE]
alldir <- rep("downregulated with age",195)
alldir[sign(codeset$log2FoldChange) <0] <- "upregulated with age"

minf <- data.frame(isgood= nmd$p_value <0.05 & nmd$samedir == TRUE, alldir, name=codeset$Customer_Identifier, in61 = codeset$Customer_Identifier %in% gm$marker, pval = nmd$p_value)


mnmd <- melt(nmd, id.vars=c("p_value", "sname", "up", "dir", "sp", "sig", "sigstar", "samedir", "YvOl2fc", "ROvOl2fc"))
mnmd$age <- "old"
mnmd$age[mnmd$variable %in% colnames(young)] <- "young"
mnmd$ing <- ""
mnmd$ing[mnmd$variable %in% gm$marker] <- "."


ggplot(mnmd, aes(x= sname, y=value, colour=age, fill=age)) +
  theme_bw()+
#  geom_point(size=3, shape=95)+
  scale_y_continuous(trans="log2")+
  facet_grid(~dir, space="free",scales="free" )+
 geom_boxplot(outlier.shape = NA)+
  scale_colour_manual(values=c("cornflower blue", "red3"))+ 
  scale_fill_manual(values=c("cornflower blue", "red3")) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=4))+
  theme(    plot.background = element_rect(fill = "transparent",colour = "transparent") )+
  xlab("Marker")+
  ylab("Expression (normalised to mean)")+
  ggtitle("All markers")

```

#Compare old vs new marker set

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

log <- sampleinfo$sex =="F" & sampleinfo$ctg %in% c(3,4,5,7) & sampleinfo$qual == "ok" & sampleinfo$reg == "" & sampleinfo$age != "N"

bcmin <- ctg1to9[,log== TRUE]
bcmin.groups <- subset(sampleinfo$type, log==TRUE)
bcmin <- subset(bcmin, rownames(bcmin) %in% minf$name[minf$in61==T])
tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups


a= ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=FALSE, labels=rownames(tbc), groups=bc.groups)+ggtitle("PCA of Old vs Young (females): 61 chosen markers")+
    theme_minimal() +
  scale_colour_manual(values=c("cornflower blue", "red3", "black"))+
  theme(plot.title=element_text(size=8,face="bold"))



bcmin <- ctg1to9[,log== TRUE]
bcmin.groups <- subset(sampleinfo$type, log==TRUE)
bcmin <- subset(bcmin, rownames(bcmin) %in% minf$name[minf$pval<0.05])
tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups

b=ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=FALSE, labels=rownames(tbc), groups=bc.groups)+ggtitle(paste("PCA of Old vs Young (females):", nrow(bcmin) ," newly chosen markers"))+
    theme_minimal() +
  scale_colour_manual(values=c("cornflower blue", "red3", "black"))+
  theme(plot.title=element_text(size=8,face="bold"))

bcmin <- ctg1to9[,log== TRUE]
bcmin.groups <- subset(sampleinfo$type, log==TRUE)

tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups

c=ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=FALSE, labels=rownames(tbc), groups=bc.groups)+ggtitle(paste("PCA of Old vs Young (females):", nrow(bcmin) ," markers"))+
    theme_minimal() +
  scale_colour_manual(values=c("cornflower blue", "red3", "black"))+
  theme(plot.title=element_text(size=8,face="bold"))


bcmin <- ctg1to9[,log== TRUE]
bcmin.groups <- subset(sampleinfo$type, log==TRUE)
bcmin <- subset(bcmin, rownames(bcmin) %in% minf$name[minf$in61==T] & rownames(bcmin) %in% minf$name[minf$isgood==T])
tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups

d=ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=FALSE, labels=rownames(tbc), groups=bc.groups)+ggtitle(paste("PCA of Old vs Young (females):", nrow(bcmin) ," newly chosen markers (both lists)"))+
    theme_minimal() +
  scale_colour_manual(values=c("cornflower blue", "red3", "black"))+
  theme(plot.title=element_text(size=8,face="bold"))


bcmin <- ctg1to9[,log== TRUE]
bcmin.groups <- subset(sampleinfo$type, log==TRUE)
bcmin <- subset(bcmin, rownames(bcmin) %in% minf$name[minf$in61==T] & rownames(bcmin) %in% minf$name[minf$isgood==F])
tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups

ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=FALSE, labels=rownames(tbc), groups=bc.groups)+ggtitle(paste("PCA of Old vs Young (females):", nrow(bcmin) ," newly chosen markers (both lists)"))+
    theme_minimal() +
  scale_colour_manual(values=c("cornflower blue", "red3", "black"))+
  theme(plot.title=element_text(size=8,face="bold"))

multiplot(a,c,b,d, cols=2)


```


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

log <- sampleinfo$sex =="F" & sampleinfo$ctg %in% c(3,4,5,7) & sampleinfo$qual == "ok" & sampleinfo$reg == "" & sampleinfo$age != "N" & sampleinfo$prep== "Luke"
bcmin <- ctg1to9[,log== TRUE]
bcmin.groups <- subset(sampleinfo$type, log==TRUE)
bcmin <- subset(bcmin, rownames(bcmin) %in% minf$name[minf$in61==T])
tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups


a= ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=T, labels=rownames(tbc),varname.size=1,varname.abbrev = TRUE, groups=bc.groups)+ggtitle("PCA of Old vs Young (females): 61 chosen markers")+
    theme_minimal() +
  scale_colour_manual(values=c("cornflower blue", "red3", "black"))+
  theme(plot.title=element_text(size=8,face="bold"))

bcmin <- ctg1to9[,log== TRUE]
bcmin.groups <- subset(sampleinfo$type, log==TRUE)
bcmin <- subset(bcmin, rownames(bcmin) %in% minf$name[minf$isgood==T])
tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups

b= ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=T, varname.size=1,varname.abbrev = TRUE,labels=rownames(tbc), groups=bc.groups)+ggtitle("PCA of Old vs Young (females): 58 newly chosen markers")+
    theme_minimal() +
  scale_colour_manual(values=c("cornflower blue", "red3", "black"))+
  theme(plot.title=element_text(size=8,face="bold"))

multiplot(a,b, cols=2)


```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
gnmd <- subset(nmd, nmd$p_value <0.05 & nmd$samedir == TRUE)

a=ggplot(data=gnmd, aes(x=YvOl2fc, y=ROvOl2fc, label=sname))+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_bw()+
  xlim(-5,+5)+
  ylim(-5,5)+
 scale_colour_manual(values= c("forest green", "cornflower blue"), name="Marker", labels=c("other","61 chosen"))+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")+
#  geom_rect(inherit.aes=FALSE, xmin=-10, ymin=-10, xmax=0, ymax=0, fill="yellow", alpha=0.004)+
#  geom_rect(inherit.aes=FALSE,xmin=0, ymin=0, xmax=10, ymax=10, fill="yellow", alpha=0.004)+
  geom_point( size=1.5, colour="cornflower blue")+
  geom_abline(inherit.aes=FALSE, intercept=0, slope=1, linetype=2)+
  theme(    plot.background = element_rect(fill = "transparent",colour = "transparent") )+
  coord_fixed(ratio=1)+
  theme(strip.text=element_blank())
  



ggsave(plot=a, height=7.5,width=15,dpi=200, device="png", filename=paste("nplot.png"), limitsize = FALSE,bg = "transparent")



b=ggplot(data=gnmd, aes(x=YvOl2fc, y=0, label=sname))+
  geom_hline(yintercept=0, colour="black")+
  theme_bw()+
  xlim(-5,+5)+
  ylim(-5,5)+
 scale_colour_manual(values= c("forest green", "cornflower blue"), name="Marker", labels=c("other","61 chosen"))+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("")+
#  geom_rect(inherit.aes=FALSE, xmin=-10, ymin=-10, xmax=0, ymax=0, fill="yellow", alpha=0.004)+
#  geom_rect(inherit.aes=FALSE,xmin=0, ymin=0, xmax=10, ymax=10, fill="yellow", alpha=0.004)+
  geom_point( size=1.5, colour="cornflower blue")+
  theme(    plot.background = element_rect(fill = "transparent",colour = "transparent") )+
  coord_fixed(ratio=1)+
  theme(strip.text=element_blank())
 

ggsave(plot=b, height=7.5,width=15,dpi=200, device="png", filename=paste("nplotb.png"), limitsize = FALSE,bg = "transparent")
 

c=ggplot(data=gnmd, aes(x=YvOl2fc, y=0, label=sname))+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_bw()+
  xlim(-5,+5)+
  ylim(-5,5)+
 scale_colour_manual(values= c("forest green", "cornflower blue"), name="Marker", labels=c("other","61 chosen"))+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")+
#  geom_rect(inherit.aes=FALSE, xmin=-10, ymin=-10, xmax=0, ymax=0, fill="yellow", alpha=0.004)+
#  geom_rect(inherit.aes=FALSE,xmin=0, ymin=0, xmax=10, ymax=10, fill="yellow", alpha=0.004)+
  geom_point( size=1.5, colour="cornflower blue")+
  geom_abline(inherit.aes=FALSE, intercept=0, slope=1, linetype=2)+
  theme(    plot.background = element_rect(fill = "transparent",colour = "transparent") )+
  coord_fixed(ratio=1)+
  theme(strip.text=element_blank())


ggsave(plot=c, height=7.5,width=15,dpi=200, device="png", filename=paste("nplotc.png"), limitsize = FALSE,bg = "transparent")

```



#Nikos plot
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
md <- ctg1to9
old <- t(subset(t(md), sampleinfo$type =="O" &  sampleinfo$sex == "F" & sampleinfo$qual =="ok" &sampleinfo$ctg %in% c(3,4,5,7)))

young <- t(subset(t(md), sampleinfo$type =="Y" & sampleinfo$sex== "F" & sampleinfo$qual =="ok"&sampleinfo$ctg %in% c(3,4,5,7)))

regen <- t(subset(t(md), sampleinfo$type =="OR" & sampleinfo$qual =="ok" &sampleinfo$ctg>2))

nmd <- md/rowMeans(cbind(young, old))

nik<- data.frame(
  fullname= rownames(md), 
  sname = substr(rownames(md),1,2), 
  YvOl2fc= foldchange2logratio(foldchange(rowMeans(young),            rowMeans(old))),
  ROvOl2fc= foldchange2logratio(foldchange(rowMeans(regen), rowMeans(old))),
  in61 =rownames(md) %in% gm$marker)

snik=subset(nik, nik$in61==TRUE)

ggplot(data=nik, aes(x=YvOl2fc, y=ROvOl2fc, label=sname, colour=in61))+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_bw()+
  xlim(-6,+6)+
  ylim(-6,6)+
 scale_colour_manual(values= c("forest green", "cornflower blue"), name="Marker", labels=c("other","61 chosen"))+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")+
#  geom_rect(inherit.aes=FALSE, xmin=-10, ymin=-10, xmax=0, ymax=0, fill="yellow", alpha=0.004)+
#  geom_rect(inherit.aes=FALSE,xmin=0, ymin=0, xmax=10, ymax=10, fill="yellow", alpha=0.004)+
  geom_point( size=8, alpha=0.4)+
  geom_abline(inherit.aes=FALSE, intercept=0, slope=1, linetype=2)+
  theme(    plot.background = element_rect(fill = "transparent",colour = "transparent") )+
  geom_text()+
  coord_fixed(ratio=1)+
  facet_wrap(~in61)+
  theme(strip.text=element_blank())
  



ggsave(plot=a, height=7.5,width=15,dpi=200, device="png", filename=paste("nplot.png"), limitsize = FALSE,bg = "transparent")


```

#Variation due to Regeneration
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
md <- t(subset(t(ctg1to9), sampleinfo$sex == "F"& sampleinfo$qual == "ok"&  sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y", "OR") ))
si <- subset(sampleinfo, sampleinfo$sample %in% colnames(md))
nmd <- as.data.frame(md/rowMeans(md))
regonly <- t(subset(t(nmd), si$type == "OR"))
nmd$regstdev <- apply(regonly, 1,sd)
nmd <- subset(nmd, rownames(nmd)%in% gm$marker)
nmd$marker <- rownames(nmd)
nmd <- nmd[order(nmd$marker),]
gm <- gm[order(gm$marker),]
nmd$dir <- gm$dir[gm$dir != "housekeeping"]

nmd$smarker <- substr(nmd$marker,1,2)
nmd$smarker <- factor(nmd$smarker, levels=nmd$smarker[order(nmd$regstdev)])


mnmd <- melt(nmd, id.vars=c("marker", "dir", "regstdev", "smarker"))
mnmd$type <- "Y"
mnmd$type[mnmd$variable %in% si$sample[si$type== "OR"]] <- "OR"
mnmd$type[mnmd$variable %in% si$sample[si$type== "O"]] <- "O"


ggplot(mnmd, aes(x= smarker, y=value, colour=type)) +
  theme_bw()+
  scale_y_continuous(trans="log2")+
  facet_grid(~dir, space="free",scales="free" )+
  geom_boxplot(outlier.shape = NA)+
  scale_colour_manual(values=c("cornflower blue", "forest green", "red3" ))+ 
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=6))
#  geom_text(aes(label=variable), size=1)


```

#Reliable rejuvenation?

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
md <- ctg1to9
old <- t(subset(t(md), sampleinfo$type =="O" &  sampleinfo$sex == "F" & sampleinfo$qual =="ok" &sampleinfo$ctg %in% c(3,4,5,7)))

young <- t(subset(t(md), sampleinfo$type =="Y" & sampleinfo$sex== "F" & sampleinfo$qual =="ok"&sampleinfo$ctg %in% c(3,4,5,7)))

regen <- t(subset(t(md), sampleinfo$type =="OR" & sampleinfo$qual =="ok" &sampleinfo$ctg>2))

oy <- cbind(old, young)

t.result <- apply(oy, 1, function (x)
t.test(x[1:ncol(old)],x[ncol(old)+1:ncol(oy)]))
ovy.p_value <- unlist(lapply(t.result, function(x) x$p.value))


oreg <- cbind(old, regen)

t.result <- apply(oreg, 1, function (x)
t.test(x[1:ncol(old)],x[ncol(old)+1:ncol(oreg)]))
ovreg.p_value <- unlist(lapply(t.result, function(x) x$p.value))

nold <- old/rowMeans(md)
nyoung <- young/rowMeans(md)
nregen <- regen/rowMeans(md)

nmd <- md/rowMeans(cbind(young, old))



nik<- data.frame(
  fullname= rownames(md), 
  sname = substr(rownames(md),1,2), 
  YvOl2fc= foldchange2logratio(foldchange(rowMeans(young),            rowMeans(old))),
  ROvOl2fc= foldchange2logratio(foldchange(rowMeans(regen), rowMeans(old))),
  in61 =rownames(md) %in% gm$marker, 
  ovypval = ovy.p_value, 
  ovrpval = ovreg.p_value,
  oldsd=  apply(nold, 1,sd),
  youngsd = apply(nyoung, 1,sd),
  regsd = apply(nregen, 1,sd))

nik$ovysig <- nik$ovypval <0.05
nik$ovrsig <- nik$ovrpval <0.1

ggplot(data=nik, aes(x=YvOl2fc, y=ROvOl2fc, label=sname, fill=regsd, shape=ovrsig))+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_minimal()+
  xlim(-6,+6)+
  ylim(-6,6)+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")+
#  geom_rect(inherit.aes=FALSE, xmin=-10, ymin=-10, xmax=0, ymax=0, fill="yellow", alpha=0.004)+
#  geom_rect(inherit.aes=FALSE,xmin=0, ymin=0, xmax=10, ymax=10, fill="yellow", alpha=0.004)+
  geom_point( size=3)+
  scale_shape_manual(values=c(21,23), labels=c(">0.1", "<0.1"), breaks=c(FALSE, TRUE) ,name="RO vs O p-value" )+
  geom_abline(inherit.aes=FALSE, intercept=0, slope=1, linetype=2)+
  theme(    plot.background = element_rect(fill = "transparent",colour = "transparent") )+
  coord_fixed(ratio=1)+
scale_fill_gradient2(high="red3", mid="white", low="blue", midpoint=4, name="Stdev (regen)")+
  facet_wrap(~ovysig)


nik$good <- nik$ROvOl2fc/nik$YvOl2fc > 0.6 & nik$ovysig ==TRUE &nik$regsd <3
nik$reju <- " neither"
nik$reju[nik$ROvOl2fc/nik$YvOl2fc > 0.6 & nik$ovysig ==TRUE &nik$regsd <3] <- "rejuvenated"
nik$reju[nik$ROvOl2fc/nik$YvOl2fc < -0.6 & nik$ovysig ==TRUE &nik$regsd <3] <- "aged"

ggplot(data=nik, aes(x=YvOl2fc, y=ROvOl2fc, label=sname, fill=regsd,colour=reju, shape=ovrsig))+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_minimal()+
  xlim(-6,+6)+
  ylim(-6,6)+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")+
#  geom_rect(inherit.aes=FALSE, xmin=-10, ymin=-10, xmax=0, ymax=0, fill="yellow", alpha=0.004)+
#  geom_rect(inherit.aes=FALSE,xmin=0, ymin=0, xmax=10, ymax=10, fill="yellow", alpha=0.004)+
  geom_point( size=3.5)+
  scale_shape_manual(values=c(21,23), labels=c(">0.1", "<0.1"), breaks=c(FALSE, TRUE) ,name="Regen vs Old \np-value" )+
  geom_abline(inherit.aes=FALSE, intercept=0, slope=1, linetype=2)+
  theme(    plot.background = element_rect(fill = "transparent",colour = "transparent") )+
  coord_fixed(ratio=1)+
scale_fill_gradient2(high="darkgoldenrod1", mid="white", low="blue", midpoint=2.5, name="Standard deviation \n amongst regenerated")+
  facet_wrap(~ovysig)+
  geom_text(size=1.7, aes(alpha=reju))+
  scale_colour_manual( values=c("red2",   "black", "dark green"), name="Rejuvenation")+
  scale_alpha_manual(values=c(1,0,1))




```


#Rejuventating Markers
Only those with:
Ratio of YvO l2fc / RO vs O l2fc > 0.6
Stdev between regenerated <3

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
md <- t(subset(t(ctg1to9), sampleinfo$sex == "F"& sampleinfo$qual == "ok"&  sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y", "OR") ))
si <- subset(sampleinfo, sampleinfo$sample %in% colnames(md))
nmd <- as.data.frame(md/rowMeans(md))
regonly <- t(subset(t(nmd), si$type == "OR"))
nmd$regstdev <- apply(regonly, 1,sd)

nmd$marker <- rownames(nmd)
nmd <- nmd[order(nmd$marker),]
codeset <- codeset[order(codeset$shortname),]
nmd$dir <- "upregulated with age"
nmd$dir[codeset$log2FoldChange > 0] <- "downregulated with age" 



nmd$smarker <- substr(nmd$marker,1,2)
nmd$smarker <- factor(nmd$smarker, levels=nmd$smarker[order(nmd$regstdev)])
nmd <- subset(nmd, nmd$marker %in% nik$fullname[nik$good ==T])

mnmd <- melt(nmd, id.vars=c("marker", "regstdev", "dir", "smarker"))
mnmd$type <- "Y"
mnmd$type[mnmd$variable %in% si$sample[si$type== "OR"]] <- "OR"
mnmd$type[mnmd$variable %in% si$sample[si$type== "O"]] <- "O"

mcodeset <- subset(codeset, codeset$Customer_Identifier %in% nmd$marker)

print(paste(mcodeset$shortname, ": " ,mcodeset$DrosoTopHit, sep=""))

print(paste(mcodeset$shortname, ": " ,mcodeset$TopHit, sep=""))

mnmd$topblast <- ""

for (i in 1:length(codeset$baseMean)){
  x = mnmd$marker == codeset$Customer_Identifier[i]
  mnmd$topblast[x] <- as.character(codeset$TopHit[i])
  mnmd$dtopblast[x] <- as.character(codeset$DrosoTopHit[i])
}



```



```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

ggplot(mnmd, aes(x= smarker, y=value, fill=type)) +
  theme_minimal()+
  scale_y_continuous(trans="log2")+
  facet_grid(~dir, space="free",scales="free" )+
#  stat_summary()+
  geom_boxplot(outlier.shape = NA)+
#  geom_point(shape=21)+
  scale_fill_manual(values=c("cornflower blue", "forest green", "red3" ))+ 
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=10))+
  xlab("Marker")+
  ylab("Expression level (normalised)")
  
#  geom_text(aes(label=variable), size=1)
mnmd$smarkerdir <- paste(substr(mnmd$dir,1,2), mnmd$smarker)
umnmd <- subset(mnmd, !duplicated(mnmd$marker))



#mcodeset$smarkerdir <- paste(substr(mcodeset$dir,1,2), mnmd$smarker)
  
  
ggplot(mnmd, aes(x= value, fill=type)) +

  theme_minimal()+
  scale_x_continuous(trans="log2", breaks=c(0.125,0.25,0.5,1,2,4), limits=c(0.125,4))+
  facet_wrap(~smarkerdir, ncol=3, scales="free_y")+
  geom_density(alpha=0.5)  +
  scale_fill_manual(values=c("cornflower blue", "forest green", "red3" ))+ 
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=10))+
  xlab("Expression level (normalised)")+
  geom_text(data=umnmd, aes(x=0.2, y=0.6, label=dtopblast), size=2,inherit.aes = F)



```



