---
title: "NanoStringNov17"
author: "Luke Hayden"
date: "24 November 2017"
output: html_document
---

##Setup
###Packages
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
#source("https://bioconductor.org/biocLite.R")
#biocLite("DESeq2")
#source("https://bioconductor.org/biocLite.R")
#biocLite("DESeq2")

library(ggplot2)
library(rmarkdown)
library(gplots)
library(RColorBrewer)
library(reshape2)
library(scales)
library(ggbiplot)
library(devtools)
library(NanoStringNorm)
library(gtools)

multiplot <- function(..., plotlist = NULL, file, cols = 1, layout = NULL) {
  require(grid)
  
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  if (is.null(layout)) {
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots == 1) {
    print(plots[[1]])
    
  } else {
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    for (i in 1:numPlots) {
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

###Data Import
We will look at our whole dataset: 108 RNA samples from 3 separate periods. 

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}


gm <- read.csv("gmark.csv")

setwd("~/Documents/nstringnov17/ctg5and6")
ct5and6.raw <- read.markup.RCC()


setwd("~/Documents/nstringnov17/ctg5hi")
ct5and6hi.raw <- read.markup.RCC()


setwd("~/Documents/nstringnov17/ctg7and8")
ct7and8.raw <- read.markup.RCC()


setwd("~/Documents/nstringnov17/ctg7and8hi")
ct7and8hi.raw <- read.markup.RCC()

setwd("~/Documents/nstringnov17/ctg3and4")

ct3and4.raw <- read.markup.RCC()


setwd("~/Documents/nstringnov17/ctg1and2")

ct1and2.raw <- read.markup.RCC()


setwd("~/Documents/nstringnov17/ctg9")

ct9.raw <- read.markup.RCC()

````

###Normalisation
3 ways of normalising data possible:
-none (raw counts)
-by internal controls 
-by internal controls and housekeeping genes

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}
codeset <- read.table("CodeSet.txt")
codeset$shortname <- substr(codeset$Customer_Identifier, 1,2)
Cht <- read.table("CodeSetCountdata.txt")

ctg1and2.norm <- NanoStringNorm(x = ct1and2.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg3and4.norm <- NanoStringNorm(x = ct3and4.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg5and6.norm <- NanoStringNorm(x = ct5and6.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg5and6hi.norm <- NanoStringNorm(x = ct5and6hi.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg7and8.norm <- NanoStringNorm(x = ct7and8.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg7and8hi.norm <- NanoStringNorm(x = ct7and8hi.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);


ctg9.norm <- NanoStringNorm(x = ct9.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);



````

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}

ctg5and6.intonly <- NanoStringNorm(x = ct5and6.raw, anno = NA, CodeCount ='sum', Background ="mean",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg5and6hi.intonly <- NanoStringNorm(x = ct5and6hi.raw, anno = NA, CodeCount ='sum', Background ="mean",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg7and8.intonly <- NanoStringNorm(x = ct7and8.raw, anno = NA, CodeCount ='sum', Background ="mean",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg7and8hi.intonly <- NanoStringNorm(x = ct7and8hi.raw, anno = NA, CodeCount ='sum', Background ="mean",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg3and4.intonly <- NanoStringNorm(x = ct3and4.raw, anno = NA, CodeCount ='sum', Background ="mean",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);


ctg1and2.intonly <- NanoStringNorm(x = ct1and2.raw, anno = NA, CodeCount ='sum', Background ="mean",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg9.intonly <- NanoStringNorm(x = ct9.raw, anno = NA, CodeCount ='sum', Background ="mean",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg1to9intonly <- cbind(ctg1and2.intonly,ctg3and4.intonly, ctg5and6hi.intonly, ctg7and8hi.intonly, ctg9.intonly)


````
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}

ctg5and6.unnorm <- NanoStringNorm(x = ct5and6.raw, anno = NA, CodeCount ='none', Background ="none",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg5and6hi.unnorm <- NanoStringNorm(x = ct5and6hi.raw, anno = NA, CodeCount ='none', Background ="none",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg7and8.unnorm <- NanoStringNorm(x = ct7and8.raw, anno = NA, CodeCount ='none', Background ="none",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg7and8hi.unnorm <- NanoStringNorm(x = ct7and8hi.raw, anno = NA, CodeCount ='none', Background ="none",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg3and4.unnorm <- NanoStringNorm(x = ct3and4.raw, anno = NA, CodeCount ='none', Background ="none",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);


ctg1and2.unnorm <- NanoStringNorm(x = ct1and2.raw, anno = NA, CodeCount ='none', Background ="none",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg9.unnorm <- NanoStringNorm(x = ct9.raw, anno = NA, CodeCount ='none', Background ="none",OtherNorm = "none", round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg1to9un <- cbind(ctg1and2.unnorm,ctg3and4.unnorm, ctg5and6hi.unnorm, ctg7and8hi.unnorm, ctg9.unnorm)


````

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}


colnames(ctg1and2.norm) <- c("C200", "C100", "C50", "C25", "C12", "C6", "CIA", "CIB", "FYA", "FOA", "MYA", "MOA","FYB",	"FOB",	"MYB",	"MOB",	"MOIA",	"MOIB","MOIC",	"MOID",	"CS1X",	"CS0.5X",	"CS0.25XA",	"CS0.25XB")
colnames(ctg1and2.unnorm) <- colnames(ctg1and2.norm) 

colnames(ctg3and4.norm) <- c("O1a", "O2a", "Y1a", "Y2a", "OR1a", "OR2a", "YR1a", "YR2a", "FY5a", "FY6a", "FO5a", "FO7a", "O1b", "O2b", "Y1b", "Y2b", "OR1b", "OR2b", "YR1b", "YR2b", "FY7", "FY8", "FO8", "FO9")
colnames(ctg3and4.unnorm) <- colnames(ctg3and4.norm) 

colnames(ctg5and6.norm) <- c("SFPA", "SFPB", "LFPA", "LFPB", "SMPA", "SMPB", "LMPA", "LMPB", "LMIA", "LMIB", "LMIC", "LMID", "Abef", "Aun", "Areg", "Bbef", "Bun", "Breg", "Cbef", "Cun", "Creg", "Dbef", "Dun", "Dreg")
colnames(ctg5and6.unnorm) <- colnames(ctg5and6.norm) 

colnames(ctg5and6hi.norm) <-  c("SFPA", "SFPB", "LFPA", "LFPB", "SMPA", "SMPB", "LMPA", "LMPB", "LMIA", "LMIB", "LMIC", "LMID", "Abef", "Aun", "Areg", "Bbef", "Bun", "Breg", "Cbef", "Cun", "Creg", "Dbef", "Dun", "Dreg")
colnames(ctg5and6hi.unnorm) <- colnames(ctg5and6hi.norm) 

colnames(ctg7and8.norm) <- c("SFPC","SFPD", "LFPC", "LFPD", "SMPC", "SMPD","LMPC", "LMPD", "0mPA", "0mPB", "10mPA", "20mPA", "Ebef", "Eun", "Ereg", "Fbef", "Fun", "Freg", "Gbef", "Gun", "Greg", "LFIA", "LFIB", "20mPB")
colnames(ctg7and8.unnorm) <- colnames(ctg7and8.norm) 

colnames(ctg7and8hi.norm) <-c("SFPC","SFPD", "LFPC", "LFPD", "SMPC", "SMPD","LMPC", "LMPD", "0mPA", "0mPB", "10mPA", "20mPA", "Ebef", "Eun", "Ereg", "Fbef", "Fun", "Freg", "Gbef", "Gun", "Greg", "LFIA", "LFIB", "20mPB")
colnames(ctg7and8hi.unnorm) <- colnames(ctg7and8hi.norm) 

colnames(ctg9.norm) <- c("LFIC", "LFID", "OmID", "10mIA", "n", "20mIC", "Ibef", "Iun", "Ireg", "Kbef", "Kun", "Kreg")
colnames(ctg9.unnorm) <- colnames(ctg9.norm) 

ctg3to9 <- cbind(ctg3and4.norm, ctg5and6hi.norm, ctg7and8hi.norm, ctg9.norm)

ctg1to9 <- cbind(ctg1and2.norm,ctg3and4.norm, ctg5and6hi.norm, ctg7and8hi.norm, ctg9.norm)

colnames(ctg1to9un) <- colnames(ctg1to9)

colnames(ctg1to9intonly) <- colnames(ctg1to9)
sampleinfo <- data.frame(

sample=colnames(ctg1to9), 
                         
sex=c(rep("M", 8), rep(c("F", "F", "M", "M"), 2), rep("M", 8), rep("F", 28), rep("M", 8), rep("F", 16), rep("M", 4), rep("F", 28)), 

age=c(rep("O",8), rep(c("Y","O","Y","O"),2),rep("O",8),  rep(c("O", "O", "Y", "Y", "O", "O", "Y", "Y","Y", "Y", "O", "O"), 2), c("Y", "Y", "O", "O", "Y", "Y", "O", "O", "O", "O", "O", "O", rep("O", 12)), c("Y", "Y", "O", "O", "Y", "Y", "O", "O", "N", "N", "N", "N"), rep("O", 11), c("N", "O", "O",  "N", "N", "N", "N", "O", "O", "O", "O", "O", "O")), 

reg=c(rep("",24), rep(c(rep("", 4), rep("R", 4), rep("", 4)), 2), c(rep("", 12), rep(c("", "", "R"),4 )   ), c(rep("", 12), rep(c("", "", "R"),3 )   ), rep("",11 ), "R", "", "", "R") , 

ctg=c(rep(1, 12), rep(2, 12),rep(3, 12), rep(4, 12), rep(5, 12), rep(6, 12), rep(7, 12), rep(8, 12), rep(9, 12)),

size= c(rep("P",6), "I", "I", rep("P",8),rep("I",4), rep("I",4), rep("P", 32), rep("I", 4), rep("P",27), rep("I", 8), "P", rep("I", 4), rep("I", 8)), 

moult = c(rep("", 80), 0,0,10,20,rep("", 11),20,"","", 0,10,"", 20,rep("", 6) ),

 exp= c(rep("conc",8), rep("OvY", 12), rep("cut", 4), rep("OvY",4), rep("regen", 4), rep("OvY",8), rep("regen", 4), rep("OvY",4), rep("OvY", 12), rep("regen", 12), rep("OvY", 8), rep("moult", 4),rep("regen", 9), "OvY", "OvY", "moult",  "OvY", "OvY", "moult", "moult", "moult", "moult", rep("regen", 6) )

)


sampleinfo$flag <- FALSE
sampleinfo$flag[c(5,6,7,8,25,31,32, 54, 59, 66,67, 78,90, 97,101)] <- TRUE


#c(23,24,37,43,44, 54, 59, 66,67, 78,90, 97,101)

sampleinfo$type <- paste(sampleinfo$age, sampleinfo$reg, sep="")
sampleinfo$qual <- "ok"
sampleinfo$qual[c(25 , 31 , 32 , 49,  54 , 67  ,76,  90, 101, 104, 107)] <- "bad"
#sampleinfo$moult <- c(rep(NA, 80), 0,0,10,20,rep(NA, 11),20,NA,NA, 0,10,NA, 20,rep(NA, 6) )

#print("Samples as follows (all experiments)")
#print(paste("males: " sum()))

```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
norm <- melt(ctg1to9)
intonly <- melt(ctg1to9intonly)
un <- melt(ctg1to9un)

normcomp <- cbind(norm, un$value, intonly$value)
colnames(normcomp) <- c("marker", "sample", "normval", "unval", "intonlyval")

p=ggplot(normcomp, aes(x=intonlyval, y=normval, colour=sample))+
  geom_point()+
  stat_smooth(model="lm")+
  scale_y_log10()+
  scale_x_log10()+
  theme(legend.position = "none")

```


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
norm <- melt(ctg1to9[,1:6])
intonly <- melt(ctg1to9intonly[,1:6])
un <- melt(ctg1to9un[,1:6])

normcomp <- cbind(norm, un$value, intonly$value)
colnames(normcomp) <- c("marker", "sample", "normval", "unval", "intonlyval")
mnormcomp <- melt(normcomp, id.vars=c("marker", "sample"))

ggplot(mnormcomp, aes(x=sample, y=value, colour=variable))+
  geom_point(position=position_dodge(width=0.2))+
  scale_y_log10()
  scale_x_log10()
  theme(legend.position = "none")

```
#Does normalisation differentially affect old vs young?
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
norm <- melt(t(subset(t(ctg1to9), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$exp == "OvY" & sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y")& sampleinfo$size== "P" )))
intonly <- melt(t(subset(t(ctg1to9intonly), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$exp == "OvY" & sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y")& sampleinfo$size== "P" )))
un <-melt(t(subset(t(ctg1to9un), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$exp == "OvY" & sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y")& sampleinfo$size== "P" )))

si <- subset(sampleinfo, sampleinfo$sample %in% colnames(ctg1to9))

normcomp <- cbind(norm, un$value, intonly$value)
colnames(normcomp) <- c("marker", "sample", "normval", "unval", "intonlyval")
normcomp$age <- "old"
normcomp$age[normcomp$sample %in% si$sample[si$age=="Y"]] <- "young" 
mnormcomp <- melt(normcomp, id.vars=c("marker", "sample", "age"))

ggplot(mnormcomp, aes(x=sample, y=value, colour=variable))+
  geom_point(position=position_dodge(width=0.4), size=1.5)+
  scale_y_log10()+
  scale_fill_manual(values=c("white", "black"))+
  facet_wrap(~age, scales="free")+
  scale_colour_manual(values=c("brown", "forest green", "orange"))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 60, hjust =1, size=7))+
  xlab("Sample")+
  ylab("Expression level (counts)")

```

Does not seem to be the case. 


#How does normalisation affect the PCA?

###Using all 195 markers
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
norm <- t(subset(t(ctg1to9), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$exp == "OvY" & sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y")& sampleinfo$size== "P" ))
intonly <- t(subset(t(ctg1to9intonly), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$exp == "OvY" & sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y")& sampleinfo$size== "P" ))
un <-t(subset(t(ctg1to9un), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$exp == "OvY" & sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y")& sampleinfo$size== "P" ))

si <- subset(sampleinfo, sampleinfo$sample %in% colnames(norm))

bcmin <- norm
bcmin.groups <- si$type

tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups

a=ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=TRUE,varname.size=1,varname.abbrev = TRUE)+
  theme_bw() +
  geom_point(aes(fill=bcmin.groups), size=3, shape=21)+
  ggtitle(paste("PCA of O vs Y: full normalisation (195 markers) " ))+ 
  scale_fill_manual(values=c("cornflower blue", "red3"))+
  theme(plot.title=element_text(size=8,face="bold"))+
  theme(axis.title=element_text(size=6))

  
bcmin <- intonly
bcmin.groups <- si$type

tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups

b=ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=TRUE,varname.size=1,varname.abbrev = TRUE)+
  theme_bw() +
  scale_fill_manual(values=c("cornflower blue", "red3"))+
  geom_point(aes(fill=bcmin.groups), size=3, shape=21)+
  ggtitle(paste("PCA of O vs Y: normalisation by internal controls only (195 markers)" ))+ 
  theme(plot.title=element_text(size=8,face="bold"))+
  theme(axis.title=element_text(size=6))

bcmin <- un
bcmin.groups <- si$type

tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups

c=ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=TRUE,varname.size=1,varname.abbrev = TRUE)+
  theme_bw() +
  scale_fill_manual(values=c("cornflower blue", "red3"))+
  geom_point(aes(fill=bcmin.groups), size=3, shape=21)+
  ggtitle(paste("PCA of O vs Y: no normalisation (195 markers)" ))+ 
  scale_colour_manual(values=c("cornflower blue", "red3"))+
  theme(plot.title=element_text(size=8,face="bold"))+
  theme(axis.title=element_text(size=6))

multiplot(c,b,a )
````

This looks good. Note how the unnormalised samples have all the eigenvectors pulling in different directions, then become nicely distributed after normalisation. 

##Using reduced marker set:


Here we have very stringent conditions, cutting our list of markers down to just:


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
md <- t(subset(t(ctg1to9), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$exp == "OvY" & sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y")& sampleinfo$size== "P" ))
si <- subset(sampleinfo, sampleinfo$sample %in% colnames(md))
old <- t(subset(t(md), si$age == "O"))
young <- t(subset(t(md), si$age == "Y"))
si <- subset(sampleinfo, sampleinfo$sample %in% colnames(norm))

norm <- t(subset(t(ctg1to9), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$exp == "OvY" & sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y")& sampleinfo$size== "P" ))
intonly <- t(subset(t(ctg1to9intonly), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$exp == "OvY" & sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y")& sampleinfo$size== "P" ))
un <-t(subset(t(ctg1to9un), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$exp == "OvY" & sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y")& sampleinfo$size== "P" ))

oy <- cbind(old, young)
t.result <- apply(oy, 1, function (x) t.test(x[1:ncol(old)],x[ncol(old)+1:ncol(oy)]))
p_value <- unlist(lapply(t.result, function(x) x$p.value))

md <- md[order(rownames(md)),]
codeset <- codeset[order(codeset$Customer_Identifier),]

YvOl2fc <- foldchange2logratio(foldchange(rowMeans(young), rowMeans(old)))
samedir <- sign(codeset$log2FoldChange) == sign(YvOl2fc) 



bcmin <- norm
bcmin.groups <- si$type
bcmin <- subset(bcmin, p_value <0.05 & samedir & rownames(bcmin) %in% gm$marker)

print(nrow(bcmin))
```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups

a=ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=TRUE,varname.size=1,varname.abbrev = TRUE)+
  theme_bw() +
  scale_fill_manual(values=c("cornflower blue", "red3"))+
  geom_point(aes(fill=bcmin.groups), size=3, shape=21)+
  ggtitle(paste("PCA of O vs Y: full normalisation (", nrow(bcmin), " markers)", sep= "" ))+ 
  theme(plot.title=element_text(size=8,face="bold"))+
  theme(axis.title=element_text(size=6))


bcmin <- intonly
bcmin.groups <- si$type

bcmin <- subset(bcmin, p_value <0.05 & samedir & rownames(bcmin) %in% gm$marker)

tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups

b=ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=TRUE,varname.size=1,varname.abbrev = TRUE)+
  theme_bw() +
  scale_fill_manual(values=c("cornflower blue", "red3"))+
  geom_point(aes(fill=bcmin.groups), size=3, shape=21)+
  ggtitle(paste("PCA of O vs Y: normalisation by internal controls only (", nrow(bcmin), " markers)", sep= "" ))+ 
  theme(plot.title=element_text(size=8,face="bold"))+
  theme(axis.title=element_text(size=6))

bcmin <- un
bcmin.groups <- si$type

bcmin <- subset(bcmin, p_value <0.05 & samedir & rownames(bcmin) %in% gm$marker)

tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups

c=ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=TRUE,varname.size=1,varname.abbrev = TRUE)+
  theme_bw() +
  scale_fill_manual(values=c("cornflower blue", "red3"))+
  geom_point(aes(fill=bcmin.groups), size=3, shape=21)+
  ggtitle(paste("PCA of O vs Y: no normalisation (", nrow(bcmin), " markers)", sep= "" ))+ 
  theme(plot.title=element_text(size=8,face="bold"))+
  theme(axis.title=element_text(size=6))

multiplot(c,b,a )
```

#How do the housekeeping genes behave?


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}

ctg1and2.whk <- NanoStringNorm(x = ct1and2.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = FALSE);
ctg1and2.whk.norm <- ctg1and2.whk$normalized.data
ctg1and2.whk.raw <- ctg1and2.whk$raw.data

ctg3and4.whk <- NanoStringNorm(x = ct3and4.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = FALSE);
ctg3and4.whk.norm <- ctg3and4.whk$normalized.data
ctg3and4.whk.raw <- ctg3and4.whk$raw.data


ctg5and6.whk <- NanoStringNorm(x = ct5and6.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = FALSE);
ctg5and6.whk.norm <- ctg5and6.whk$normalized.data
ctg5and6.whk.raw <- ctg5and6.whk$raw.data


ctg5and6hi.whk <- NanoStringNorm(x = ct5and6hi.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = FALSE);
ctg5and6hi.whk.norm <- ctg5and6hi.whk$normalized.data
ctg5and6hi.whk.raw <- ctg5and6hi.whk$raw.data


ctg7and8.whk <- NanoStringNorm(x = ct7and8.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = FALSE);
ctg7and8.whk.norm <- ctg7and8.whk$normalized.data
ctg7and8.whk.raw <- ctg7and8.whk$raw.data


ctg7and8hi.whk <- NanoStringNorm(x = ct7and8hi.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = FALSE);
ctg7and8hi.whk.norm <- ctg7and8hi.whk$normalized.data
ctg7and8hi.whk.raw <- ctg7and8hi.whk$raw.data


ctg9.whk <- NanoStringNorm(x = ct9.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = FALSE);
ctg9.whk.norm <- ctg9.whk$normalized.data
ctg9.whk.raw <- ctg9.whk$raw.data

ctg1to9.whk.norm <- cbind(ctg1and2.whk.norm,ctg3and4.whk.norm[,4:27], ctg5and6hi.whk.norm[,4:27], ctg7and8hi.whk.norm[,4:27], ctg9.whk.norm[,4:15])

colnames(ctg1to9.whk.norm) <- c(colnames(ctg1and2.whk.norm[,1:3]), colnames(ctg1to9))

ctg1to9.whk.raw <- cbind(ctg1and2.whk.raw,ctg3and4.whk.raw[,4:27], ctg5and6hi.whk.raw[,4:27], ctg7and8hi.whk.raw[,4:27], ctg9.whk.raw[,4:15])

colnames(ctg1to9.whk.raw) <- c(colnames(ctg1and2.whk.raw[,1:3]), colnames(ctg1to9))

hk.norm <- subset(ctg1to9.whk.norm, ctg1to9.whk.norm[,1] == "Housekeeping")

hk.raw <- subset(ctg1to9.whk.raw, ctg1to9.whk.raw[,1] == "Housekeeping")
```



```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
norm <- hk.norm[,4:111]

norm <- t(subset(t(norm), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$exp == "OvY" & sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y")& sampleinfo$size== "P" ))

si <- subset(sampleinfo, sampleinfo$sample %in% colnames(norm))

bcmin <- norm
bcmin.groups <- si$type

tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups

ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=TRUE,varname.size=1,varname.abbrev = TRUE)+
  theme_bw() +
  scale_fill_manual(values=c("cornflower blue", "red3"))+
  geom_point(aes(fill=bcmin.groups), size=3, shape=21)+
  ggtitle(paste("PCA of old vs young using housekeeping genes values (normalised)", sep= "" ))+ 
  theme(plot.title=element_text(size=8,face="bold"))+
  theme(axis.title=element_text(size=6))


raw <- hk.raw[,4:111]

raw<- t(subset(t(raw), sampleinfo$sex == "F"& sampleinfo$qual == "ok"& sampleinfo$exp == "OvY" & sampleinfo$ctg>2& sampleinfo$type %in% c("O", "Y")& sampleinfo$size== "P" ))

si <- subset(sampleinfo, sampleinfo$sample %in% colnames(raw))

bcmin <- raw
bcmin.groups <- si$type

tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups

ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=TRUE,varname.size=1,varname.abbrev = TRUE)+
  theme_bw() +
  scale_fill_manual(values=c("cornflower blue", "red3"))+
  geom_point(aes(fill=bcmin.groups), size=3, shape=21)+
  ggtitle(paste("PCA of old vs young using housekeeping genes values (raw)", sep= "" ))+ 
  theme(plot.title=element_text(size=8,face="bold"))+
  theme(axis.title=element_text(size=6))

```
