---
title: "NanoStringNov17"
author: "Luke Hayden"
date: "8th March 2017"
output: html_document
---



file:///C:/Users/Luke/Documents/nstringnov17/20170208_lh 8-2-17_lh 8 2 17_12.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170207_luke 7 2 17_lh 7 2 17_01.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170207_luke 7 2 17_lh 7 2 17_02.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170207_luke 7 2 17_lh 7 2 17_03.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170207_luke 7 2 17_lh 7 2 17_04.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170207_luke 7 2 17_lh 7 2 17_05.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170207_luke 7 2 17_lh 7 2 17_06.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170207_luke 7 2 17_lh 7 2 17_07.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170207_luke 7 2 17_lh 7 2 17_08.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170207_luke 7 2 17_lh 7 2 17_09.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170207_luke 7 2 17_lh 7 2 17_10.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170207_luke 7 2 17_lh 7 2 17_11.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170207_luke 7 2 17_lh 7 2 17_12.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170208_lh 8-2-17_lh 8 2 17_01.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170208_lh 8-2-17_lh 8 2 17_02.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170208_lh 8-2-17_lh 8 2 17_03.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170208_lh 8-2-17_lh 8 2 17_04.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170208_lh 8-2-17_lh 8 2 17_05.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170208_lh 8-2-17_lh 8 2 17_06.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170208_lh 8-2-17_lh 8 2 17_07.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170208_lh 8-2-17_lh 8 2 17_08.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170208_lh 8-2-17_lh 8 2 17_09.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170208_lh 8-2-17_lh 8 2 17_10.RCC
file:///C:/Users/Luke/Documents/nstringnov17/20170208_lh 8-2-17_lh 8 2 17_11.RCC

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
#source("https://bioconductor.org/biocLite.R")
#biocLite("DESeq2")
#source("https://bioconductor.org/biocLite.R")
#biocLite("DESeq2")

library(ggplot2)
library(DESeq2)
library(rmarkdown)
library(pracma)
library(gplots)
library(RColorBrewer)
library(reshape2)
library(scales)
library(ggbiplot)
library(gtools)
library(devtools)
library(NanoStringNorm)
library(ggthemes)
library(grid)
library(gridExtra)
library(sjstats)


multiplot <- function(..., plotlist = NULL, file, cols = 1, layout = NULL) {
  require(grid)
  
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  if (is.null(layout)) {
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots == 1) {
    print(plots[[1]])
    
  } else {
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    for (i in 1:numPlots) {
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```



```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}


gm <- read.csv("gmark.csv")

setwd("~/nstringnov17/ctg5and6")
ct5and6.raw <- read.markup.RCC()


setwd("~/nstringnov17/ctg5hi")
ct5and6hi.raw <- read.markup.RCC()


setwd("~/nstringnov17/ctg7and8")
ct7and8.raw <- read.markup.RCC()

setwd("~/nstringnov17/ctg3and4")

ct3and4.raw <- read.markup.RCC()

````

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}
codeset <- read.table("CodeSet.txt")
codeset$shortname <- substr(codeset$Customer_Identifier, 1,2)
Cht <- read.table("CodeSetCountdata.txt")




ctg5and6.norm <- NanoStringNorm(x = ct5and6.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);


ctg5and6hi.norm <- NanoStringNorm(x = ct5and6hi.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg7and8.norm <- NanoStringNorm(x = ct7and8.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

#ctg7and8hi.norm <- NanoStringNorm(x = ct7and8hi.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);

ctg3and4.norm <- NanoStringNorm(x = ct3and4.raw, anno = NA, CodeCount ='sum', Background ="mean", SampleContent ='housekeeping.sum', round.values = FALSE, take.log = FALSE,return.matrix.of.endogenous.probes = TRUE);
3and4.raw <- read.markup.RCC()

````
#Normalisation
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}

colnames(ctg5and6.norm) <- c("SFPA", "SFPB", "LFPA", "LFPB", "SMPA", "SMPB", "LMPA", "LMPB", "LMIA", "LMIB", "LMIC", "LMID", "Abef", "Aun", "Areg", "Bbef", "Bun", "Breg", "Cbef", "Cun", "Creg", "Dbef", "Dun", "Dreg")

colnames(ctg5and6hi.norm) <-  c("SFPA", "SFPB", "LFPA", "LFPB", "SMPA", "SMPB", "LMPA", "LMPB", "LMIA", "LMIB", "LMIC", "LMID", "Abef", "Aun", "Areg", "Bbef", "Bun", "Breg", "Cbef", "Cun", "Creg", "Dbef", "Dun", "Dreg")


colnames(ctg7and8.norm) <- c("SFPC","SFPD", "LFPC", "LFPD", "SMPC", "SMPD","LMPC", "LMPD", "0mPA", "0mPB", "10mPA", "20mPA", "Ebef", "Eun", "Ereg", "Fbef", "Fun", "Freg", "Gbef", "Gun", "Greg", "LFIA", "LFIB", "20mPB")

#colnames(ctg7and8hi.norm) <-c("SFPC","SFPD", "LFPC", "LFPD", "SMPC", "SMPD","LMPC", "LMPD", "0mPA", "0mPB", "10mPA", "20mPA", "Ebef", "Eun", "Ereg", "Fbef", "Fun", "Freg", "Gbef", "Gun", "Greg", "LFIA", "LFIB", "20mPB")


colnames(ctg3and4.norm) <- c("O1a", "O2a", "Y1a", "Y2a", "OR1a", "OR2a", "YR1a", "YR2a", "FY5a", "FY6a", "FO5a", "FO7a", "O1b", "O2b", "Y1b", "Y2b", "OR1b", "OR2b", "YR1b", "YR2b", "FY7", "FY8", "FO8", "FO9")

ctg3to8 <- cbind(ctg3and4.norm, ctg5and6hi.norm, ctg7and8.norm)
sampleinfo <- data.frame(sample=colnames(ctg3to8), sex=c(rep("F", 28), rep("M", 8), rep("F", 16), rep("M", 8), rep("F", 12)), age=c(rep(c("O", "O", "Y", "Y", "O", "O", "Y", "Y","Y", "Y", "O", "O"), 2), rep(c("Y", "Y", "O", "O", "Y", "Y", "O", "O", "O", "O", "O", "O", rep("O", 12)), 2)), reg=c(rep(c(rep("", 4), rep("R", 4), rep("", 4)), 2), c(rep("", 12), rep(c("", "", "R"),4 )   ), c(rep("", 12), rep(c("", "", "R"),3 )   ), "", "", "") , ctg=c(rep("3", 12), rep("4", 12), rep("5", 12), rep("6", 12), rep("7", 12), rep("8", 12))  )
sampleinfo$type <- paste(sampleinfo$age, sampleinfo$reg, sep="")
sampleinfo$qual <- "ok"
sampleinfo$qual[c(1,7,8, 25, 43, 52, 66)] <- "bad"

```
Sample Quality

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}
#all <- ctg3to8
#all$isg <- rownames(ctg3to8) %in% gm
mc36 <- melt(ctg3to8)
mc36$exp <- mc36$value > 1
mc36$isg <- mc36$Var1 %in% gm$marker
ggplot(mc36, aes(y=value, x=Var1, colour=isg)) +geom_point(alpha=0.3) +facet_wrap(~Var2) +scale_y_log10() 


#ggplot(mc36, aes(y=value, x=Var1, colour=exp)) +geom_point(alpha=0.3, shape=95, size=5) +facet_wrap(~Var2) +scale_y_log10() + theme(  axis.text.x = element_blank())

```


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
mc5 <- melt(ctg5and6.norm)
mc5h <- melt(ctg5and6hi.norm)
mc5h$hival <- mc5$value

ggplot(data=mc5h, aes(x=value, y=hival, colour=Var2)) +geom_point() +facet_wrap(~Var2, ncol=6) +scale_x_log10() +scale_y_log10()
```



```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
#bcmin <- all.norm[,c(1:4)]
ctg5h.norm <- ctg5hi.norm
colnames(ctg5h.norm) <- paste(colnames(ctg5hi.norm),"h")
bcmin <- cbind(ctg5.norm, ctg5h.norm)
tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- c(rep(c("Y", "Y", "O","O","Y", "Y", "O","O","O","O","O","O"),2))
#bc.groups <- c(rep("low",12), rep("high", 12))

#bc.groups <- c("Y", "Y", "O","O")

ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=FALSE, labels=rownames(tbc), groups=bc.groups)+ theme(legend.direction = 'horizontal', legend.position = 'none')+ggtitle("PCA of O vs Y: 195 markers")+ scale_colour_manual(values=c("cornflower blue", "red3"))

```



```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

bcmin <- ctg3and4.norm[,c(2,3,4,9,10,11,12,13:16,21:24)]
bcmin <- subset(bcmin, rownames(bcmin) %in% gm$marker)
tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- c("O", "Y", "Y", "Y", "Y", "O", "O", "O", "O", "Y", "Y", "Y", "Y", "O", "O")

ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=FALSE, labels=rownames(tbc), groups=bc.groups)+ theme(legend.direction = 'horizontal', legend.position = 'none')+ggtitle("PCA of O vs Y")+ scale_colour_manual(values=c("cornflower blue", "red3"))


rg <-subset(ctg5.norm, rownames(ctg5.norm) %in% gm$marker)

rg.sc <- scale(t(rg), center= bc.pca$center)
rg.pred <- rg.sc %*% bc.pca$rotation 


rgd.pca <- bc.pca
rgd.pca$x <- rbind(bc.pca$x, rg.pred)
                   
                   
rgd.groups <- c("O", "Y", "Y", "Y", "Y", "O", "O", "O", "O", "Y", "Y", "Y", "Y", "O", "O", "Y", "Y", "O","O","Y", "Y", "O","O","O","O","O","O")                 

ggbiplot(rgd.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=FALSE, labels=rownames(rgd.pca$x), groups=rgd.groups, varname.abbrev = TRUE)+ theme(legend.direction = 'horizontal', legend.position = 'top')+ggtitle("PCA Old vs Young, using just our subset of 61 markers + new projected")+ scale_colour_manual(values=c("cornflower blue","red3"))


```


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}




bcmin <- ctg3to8[,c(2,3,4,9,10,11,12,13:16,21:24,26:28,49:52)]
bcmin <- subset(bcmin, rownames(bcmin) %in% gm$marker)
tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- c( "O", "Y", "Y", "Y", "Y", "O", "O", "O", "O", "Y", "Y", "Y", "Y", "O", "O","Y","O","O","Y", "Y","O","O")

ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=FALSE, labels=rownames(tbc), groups=bc.groups)+ theme(legend.direction = 'horizontal', legend.position = 'none')+ggtitle("PCA of O vs Y: 61 chosen markers")+ scale_colour_manual(values=c("cornflower blue", "red3"))



````


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

log <- sampleinfo$sex =="F" & sampleinfo$ctg %in% c(3,4,5,7) & sampleinfo$qual == "ok" & sampleinfo$reg == ""


bcmin <- ctg3to8[,log== TRUE]
bcmin.groups <- subset(sampleinfo$type, log==TRUE)
bcmin <- subset(bcmin, rownames(bcmin) %in% gm$marker)
tbc <- t(bcmin)
tbc <- tbc[,colSums(tbc) > 0]
bc.pca <- prcomp(tbc,center = TRUE,scale. = TRUE) 
bc.groups <- bcmin.groups

ggbiplot(bc.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=FALSE, labels=rownames(tbc), groups=bc.groups)+ theme(legend.direction = 'horizontal', legend.position = 'none')+ggtitle("PCA of O vs Y: 61 chosen markers")+ scale_colour_manual(values=c("cornflower blue", "red3"))



rg <-ctg3to8[,c(5,6,17,18,37:42,44:48,61:69)]
rg <- subset(rg, rownames(rg) %in% gm$marker)
rg.groups <- sampleinfo$type[c(5,6,17,18,37:42,44:48,61:69)]

rg.sc <- scale(t(rg), center= bc.pca$center)
rg.pred <- rg.sc %*% bc.pca$rotation 


rgd.pca <- bc.pca
rgd.pca$x <- rbind(bc.pca$x, rg.pred)
                   
                   
rgd.groups <- append(bcmin.groups, rg.groups)                

ggbiplot(rgd.pca, obs.scale = 1, var.scale = 1, ellipse = FALSE, circle = TRUE, var.axes=FALSE, labels=rownames(rgd.pca$x), groups=rgd.groups, varname.abbrev = TRUE)+ theme(legend.direction = 'horizontal', legend.position = 'top')+ggtitle("PCA Old vs Young, using just our subset of 61 markers + new projected")+ scale_colour_manual(values=c("cornflower blue", "forest green", "red3"))

````


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
an <- ctg3to8/rowMeans(ctg3to8)
an <- subset(an, rownames(an) %in% gm$marker)
log <- sampleinfo$sex =="F" &  sampleinfo$qual == "ok" & sampleinfo$type != "YR"
anf <- an[,log==TRUE]

sif <- subset(sampleinfo, log)
colnames(anf) <- paste( sif$type, colnames(anf))
#anf <- rbind(anf, t(sif))
manf <- melt(anf)

manf$type <- substr(manf$Var2,1,2)
manf$mk <- substr(manf$Var1, 1,2)
manf$dir <- "upregulated with age"
manf$dir[manf$Var1 %in% gm$marker[gm$dir =="downregulated with age"]] <- "downregulated with age"
manf$dm <- paste(substr(manf$dir, 1,1), manf$mk)

ggplot(manf, aes(x=mk, y=value, colour=type)) +geom_point(shape=95, size=4)  +scale_color_manual(values=c("cornflower blue", "red3", "forest green"), name= "Age", labels=c("Old", "Young", "Regenerated Old")) +scale_y_log10()


ggplot(manf, aes(x=mk, y=value, colour=type, fill=dir)) +
  geom_boxplot()+
  scale_y_continuous(trans='log2', breaks=c(0,0.25,0.5,1,2,4))+
  geom_hline(yintercept=1, colour = "black", linetype=2) +
  scale_fill_manual(values=c("white", "dark grey"))  +
  scale_color_manual(values=c("cornflower blue", "forest green",  "red3")) +
  theme(axis.text.x = element_text(angle = 60, hjust =0.5, size=7)) +
  facet_wrap(~dir, ncol=1, scales="free_x")


ggplot(manf, aes(x=value, fill=type)) +
  geom_density(alpha=0.5)  +
  scale_fill_manual(values=c("cornflower blue",  "red3", "forest green"), name= "Age", labels=c("Old", "Young", "Regenerated Old")) +
  scale_x_log10() +
  facet_wrap(~dir, ncol=1  )

````

#Regeneration exp in detail
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}



````
#Regeneration exp in detail
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
an <- ctg3to8/rowMeans(ctg3to8)
an <-subset(an, rownames(an) %in% gm$marker)
re <- an[,c(37:48,61:69)]
msi <- subset(sampleinfo, sampleinfo$sample %in% colnames(re))
re <- re[, msi$qual=="ok"]
                


mre <- melt(re)
mre$sample <- substr(mre$Var2,1,1)
mre$type <- substr(mre$Var2,2,length(mre$Var2))
mre$mk <- substr(mre$Var1, 1,2)
mre$dir <- "upregulated with age"
mre$dir[mre$Var1 %in% gm$marker[gm$dir =="downregulated with age"]] <- "downregulated with age"
mre$md <- paste(substr(mre$dir, 1,2), mre$mk)

ggplot(data=mre, aes(colour=type, x=md, y=value)) +
  geom_point() +
  scale_y_continuous(trans='log2', breaks=c(0,0.25,0.5,1,2,4))+
  geom_hline(yintercept=1, colour = "black", linetype=2)+
#  facet_wrap(~sample, ncol=1) +
  scale_color_manual(values=c("cornflower blue", "forest green",  "dark blue")) +
  theme(axis.text.x = element_text(angle = 60, hjust=1))+
  stat_summary_bin()

````

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
ggplot(data=mre, aes(colour=type, x=md, y=value, fill=dir)) +
  geom_boxplot(outlier.shape=NA) +
  scale_y_continuous(trans='log2', breaks=c(0,0.25,0.5,1,2,4))+
  geom_hline(yintercept=1, colour = "black", linetype=2)+
#  facet_wrap(~sample, ncol=1) +
  scale_color_manual(values=c("cornflower blue", "forest green",  "dark blue")) +
  theme(axis.text.x = element_text(angle = 60, hjust=1))+
  scale_fill_manual(values=c("white", "dark grey")) 


ggplot(data=mre, aes(colour=type, x=md, y=value)) +
  geom_point() +
  scale_y_continuous(trans='log2', breaks=c(0,0.25,0.5,1,2,4))+
  geom_hline(yintercept=1, colour = "black", linetype=2)+
#  facet_wrap(~sample, ncol=1) +
  scale_color_manual(values=c("cornflower blue", "forest green",  "dark blue")) +
  theme(axis.text.x = element_text(angle = 60, hjust=1))+
  facet_wrap(~sample)



ggplot(data=mre, aes(fill=type, x=value)) +
  geom_density(alpha=0.5) +
  scale_x_continuous(trans='log2', breaks=c(0,0.25,0.5,1,2,4))+
  geom_vline(xintercept=1, colour = "black", linetype=2)+
  scale_fill_manual(values=c("cornflower blue", "forest green",  "dark blue")) +
  facet_wrap(~dir, ncol=1)

````



