---
title: "Rejuvenating and Hyper-aging markers"
author: "Luke Hayden"
date: "April 12, 2018"
output: html_document
---

#Setup
###Packages
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
#source("https://bioconductor.org/biocLite.R")
#biocLite("DESeq2")
#source("https://bioconductor.org/biocLite.R")
#biocLite("DESeq2")

library(ggplot2)
library(rmarkdown)
library(gplots)
library(RColorBrewer)
library(reshape2)
library(scales)
library(ggbiplot)
library(devtools)
library(NanoStringNorm)
library(gtools)
library(Rmisc)
library(dplyr)
library(tidyr)
library(caret)
library(e1071)
library(caTools)
library(FinCal)

```

###Data Import

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}
setwd("~/Documents/nstringnov17")
load(file="nov17main_data.rdata")
```

#Nikos plot

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
md <- ctg1to9

old <- t(subset(t(md), sampleinfo$type =="O" &  sampleinfo$sex == "F" & sampleinfo$qual =="ok" &sampleinfo$ctg %in% c(3,4,5,7)))

young <- t(subset(t(md), sampleinfo$type =="Y" & sampleinfo$sex== "F" & sampleinfo$qual =="ok"&sampleinfo$ctg %in% c(3,4,5,7)))

regen <- t(subset(t(md), sampleinfo$type =="OR" & sampleinfo$qual =="ok" &sampleinfo$ctg>2))

vardat <- data.frame(name=rownames(old), 
                     oldmean=rowMeans(old), 
                     oldsd= apply(old,1, sd), 
                     oldcv = coefficient.variation(rowMeans(old), apply(old,1, sd)),
                   youngmean=rowMeans(young), 
                     youngsd= apply(young,1, sd), 
                     youngcv = coefficient.variation(rowMeans(young), apply(young,1, sd)), 
                   regenmean=rowMeans(regen), 
                     regensd= apply(regen,1, sd), 
                     regencv = coefficient.variation(rowMeans(regen), apply(regen,1, sd))
                   ) %>%
  left_join(minf, by="name")


oreg <- cbind(old, regen)

t.result <- apply(oreg, 1, function (x)
t.test(x[1:ncol(old)],x[ncol(old)+1:ncol(oreg)]))
vardat$ovreg.pval <- unlist(lapply(t.result, function(x) x$p.value))
vardat$ovregsig <- vardat$ovreg.pval <0.1

ggplot(data=vardat, aes(x=YvOl2fc, y=ROvOl2fc, label=sname, colour=chosen))+
  geom_point()+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_bw()+
  xlim(-6,+6)+
  ylim(-6,6)+
  scale_colour_manual(values= c("grey", "forest green"), name="Marker", labels=c("other","chosen"))+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")


```



```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
vardat2 <- subset(vardat, vardat$chosen ==TRUE)

ggplot(data=vardat2, aes(x=YvOl2fc, y=ROvOl2fc, label=sname, fill=regencv, shape=ovregsig))+
  geom_point(size=3)+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_minimal()+
  xlim(-6,+6)+
  ylim(-6,6)+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")+
  scale_shape_manual(values=c(21,23), labels=c(">0.1", "<0.1"), breaks=c(FALSE, TRUE) ,name="RO vs O p-value" )+
  scale_fill_gradient2(high="red3", mid="white", low="blue", midpoint=0.6, name="cv (regen)")+
  geom_abline(inherit.aes=FALSE, intercept=0, slope=1, linetype=2)



```

#Pairwise log2foldchange

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
md <-t(subset(t(ctg1to9), sampleinfo$type %in% c("O", "Y", "OR") &  sampleinfo$sex == "F" & sampleinfo$qual =="ok" &sampleinfo$ctg %in% c(3,4,5,7)))

si <- subset(sampleinfo, sampleinfo$sample %in% colnames(md))

YvOpairfolds <- data.frame(matrix(ncol=0, nrow=nrow(md)))
addfolds <- data.frame(matrix(ncol=0, nrow=nrow(md)))
pairnames <- c()

for (yi in 1:ncol(young)){
  for (oi in 1:ncol(old)){
    addfolds <- cbind(addfolds, foldchange2logratio(foldchange(young[,yi], old[,oi])))
    pairnames <- append(pairnames, paste(colnames(young)[yi], colnames(old)[oi], sep="-"))
  }
  colnames(addfolds) <- pairnames
  YvOpairfolds <- cbind(YvOpairfolds, addfolds)
  
  addfolds <- data.frame(matrix(ncol=0, nrow=nrow(md)))
  pairnames <- c()

}


ROvOpairfolds <- data.frame(matrix(ncol=0, nrow=nrow(md)))


for (yi in 1:ncol(regen)){
  for (oi in 1:ncol(old)){
    addfolds <- cbind(addfolds, foldchange2logratio(foldchange(regen[,yi], old[,oi])))
    pairnames <- append(pairnames, paste(colnames(regen)[yi], colnames(old)[oi], sep="-"))
  }
  colnames(addfolds) <- pairnames
  ROvOpairfolds <- cbind(ROvOpairfolds, addfolds)
  
  addfolds <- data.frame(matrix(ncol=0, nrow=nrow(md)))
  pairnames <- c()

}

YvOpairfoldsna <- YvOpairfolds
YvOpairfoldsna[mapply(is.infinite, YvOpairfoldsna)] <- NA

ROvOpairfoldsna <- ROvOpairfolds
ROvOpairfoldsna[mapply(is.infinite, ROvOpairfoldsna)] <- NA

```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

pairfoldsum <- data.frame(
  name=rownames(YvOpairfolds), 
  YvOfcmean = rowMeans(YvOpairfoldsna, na.rm=T), 
  YvOfcsd = apply(YvOpairfoldsna, 1, sd, na.rm=T), 
  ROvOfcmean = rowMeans(ROvOpairfoldsna, na.rm=T), 
  ROvOfcsd = apply(ROvOpairfoldsna, 1, sd, na.rm=T) 
) %>%
  left_join(minf, by="name")%>%
  filter(chosen==T)

ggplot(pairfoldsum, aes())



```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}





YvOpairfolds2 <- YvOpairfolds%>% 
  mutate(name= rownames(YvOpairfolds))%>%
  subset(rownames(YvOpairfolds) %in% minf$name[minf$chosen==T]) %>%
  gather(key="sample", value="pairwise YvOl2fc", -name)




ROvOpairfolds2 <- ROvOpairfolds %>% 
  mutate(name= rownames(YvOpairfolds))%>%
  subset(rownames(ROvOpairfolds) %in% minf$name[minf$chosen==T]) %>%
  gather(key="sample", value="pairwise ROvOl2fc", -name)

YvOpairfolds2 <-  left_join(YvOpairfolds2,
                            data.frame(name=rownames(ROvOpairfolds),
                                       meanROvOlfc= rowMeans(ROvOpairfoldsna, na.rm=T)), 
                            by="name")


ROvOpairfolds2 <-  left_join(ROvOpairfolds2,
                            data.frame(name=rownames(YvOpairfolds),
                                       meanYvOlfc= rowMeans(YvOpairfoldsna, na.rm=T)), 
                            by="name")

```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}



ggplot(data=ROvOpairfolds2, aes(x= `pairwise ROvOl2fc`, y=meanYvOlfc))+
         geom_point()+
  geom_point(data=YvOpairfolds2, aes(y= `pairwise YvOl2fc`, x=meanROvOlfc))

```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

ggplot(data=vardat2, aes(y=ROvOl2fc, label=sname, x=regencv, shape=ovregsig, fill=dir))+
  geom_point(size=3)+
  theme_minimal()+
  xlab("Coefficient of variation (amongst regenerated samples)") +
 ylab("Regenerated vs Old Log2 Fold Change")+
  scale_fill_manual(values=c("dark blue", "forest green"))+
 scale_shape_manual(values=c(21,23), labels=c(">0.1", "<0.1"), breaks=c(FALSE, TRUE) ,name="RO vs O p-value" )


ggplot(vardat2, aes(x=sname, y=regencv))+geom_point()
```

#Rejuvenating Markers
Only those with:
from chosen 80 markers
Ratio of YvO l2fc / RO vs O l2fc > 0.6
cv between regenerated <0.5

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
vardat$rejuv <- "other"

cvthresh<- 0.4


vardat$rejuv[vardat$chosen== TRUE & vardat$regencv < cvthresh & vardat$ROvOl2fc/vardat$YvOl2fc > 0.6] <- "rejuvenated"


vardat$rejuv[vardat$chosen== TRUE & vardat$regencv < cvthresh & vardat$ROvOl2fc/vardat$YvOl2fc < -0.6] <- "aged"


vardat2 <- subset(vardat, vardat$chosen ==TRUE)
ggplot(data=vardat2, aes(x=YvOl2fc, y=ROvOl2fc, label=sname, fill=regencv, shape=ovregsig, colour=rejuv))+
  geom_point(size=3)+
  geom_hline(yintercept=0, colour="black")+
  geom_vline(xintercept=0, colour="black")+
  theme_minimal()+
  xlim(-6,+6)+
  ylim(-6,6)+
  xlab("Young vs Old Log2 Fold Change") +
  ylab("Regenerated vs Old Log2 Fold Change")+
  scale_shape_manual(values=c(21,23), labels=c(">0.1", "<0.1"), breaks=c(FALSE, TRUE) ,name="RO vs O p-value" )+
  scale_fill_gradient2(high="red3", mid="white", low="blue", midpoint=0.6, name="cv (regen)")+
  geom_abline(inherit.aes=FALSE, intercept=0, slope=1, linetype=2)+
  guides(fill=F)+
  scale_colour_manual(values=c("violet","black", "forest green" ))



md <- ctg1to9[,colnames(ctg1to9) %in% c(colnames(old), colnames(young), colnames(regen))]
md <- data.frame(md/rowMeans(md)) 
md$name <- rownames(md)
md <- inner_join(md, vardat2, by="name") %>% 
  gather(key="sample", value="normexp", 1:sum(ncol(old), ncol(young), ncol(regen))) %>%
  left_join(sampleinfo, by="sample")%>%  inner_join(codeset, by=c("name" = "Customer_Identifier"))

ggplot(data=subset(md, md$rejuv != "other"), aes(x=sname, y=normexp, fill=type, colour=type))+
#  geom_point(shape=95)+
  geom_violin(outlier.shape = NA)+
  geom_hline(yintercept=1, linetype=2)+
#  stat_summary()+
  facet_grid(~rejuv, space="free", scale="free")+
  scale_fill_manual(values=c("cornflower blue", "forest green", "red3"))+
  scale_colour_manual(values=c("cornflower blue", "forest green", "red3"))+
  theme_minimal()+
  scale_y_continuous(trans="log2", breaks=c(0.125,0.25,0.5,1,2,4))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=10))+
  xlab("Marker")+
  ylab("Expression level (normalised)")



ggplot(data=subset(md, md$rejuv != "other"), aes(x=sname, y=normexp, fill=type, colour=type))+
#  geom_point(shape=95)+
#  geom_violin(outlier.shape = NA)+
  geom_hline(yintercept=1, linetype=2)+
  stat_summary()+
  facet_grid(~rejuv, space="free", scale="free")+
  scale_fill_manual(values=c("cornflower blue", "forest green", "red3"))+
  scale_colour_manual(values=c("cornflower blue", "forest green", "red3"))+
  theme_minimal()+
  scale_y_continuous(trans="log2", breaks=c(0.125,0.25,0.5,1,2,4))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=10))+
  xlab("Marker")+
  ylab("Expression level (normalised)")


```



```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

ggplot(data=md, aes(x=type, y=normexp, fill=type))+
#  geom_point(shape=95)+
  geom_violin(outlier.shape = NA)+
  geom_hline(yintercept=1, linetype=2)+
#  stat_summary()+
  facet_grid(~rejuv, space="free", scale="free")+
  scale_fill_manual(values=c("cornflower blue", "forest green", "red3"))+
  theme_minimal()+
  scale_y_continuous(trans="log2", breaks=c(0.125,0.25,0.5,1,2,4))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=10))+
  xlab("Marker")+
  ylab("Expression level (normalised)")


ggplot(data=md, aes(x=type, y=normexp, colour=type))+
  geom_point(alpha=0.2)+
#  geom_violin(outlier.shape = NA)+
  geom_hline(yintercept=1, linetype=2)+
#  stat_summary()+
  facet_grid(~rejuv, space="free", scale="free")+
  scale_colour_manual(values=c("cornflower blue", "forest green", "red3"))+
  theme_minimal()+
  scale_y_continuous(trans="log2", breaks=c(0.125,0.25,0.5,1,2,4))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=10))+
  xlab("Marker")+
  ylab("Expression level (normalised)")



```


#Hyper-aged and rejuvenated marker identity

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
print("rejuvenated:")
levels(as.factor(as.character(subset(md, md$rejuv == "rejuvenated")$TopHit)))

print("aged:")
levels(as.factor(as.character(subset(md, md$rejuv == "aged")$TopHit)))

```